---
title: "Investigating Physiological Changes In Humans Induced By Fasting and Refeeding (FastForward Study)"
author: "Megat Nabil Jamaludin bin Megat Faizal"
date: "2026-02-22"
output:
  html_document:
    toc: true
    toc_depth: 3
    theme: flatly
---

```{r setup, include=FALSE}
options(encoding = "UTF-8")
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width = 18,
  fig.height = 12
)

## =========================
## PACKAGES
## =========================
library(tidyverse)   
library(lme4)
library(lmerTest)
library(emmeans)
library(broom)
library(broom.mixed)
library(kableExtra)
library(knitr)
library(readxl)
library(e1071)
library(car)
library(purrr)

## =========================
## HELPERS
## =========================
first_non_na <- function(x) {
  y <- x[!is.na(x)]
  if (length(y) == 0) NA else y[1]
}

center <- function(x) x - mean(x, na.rm = TRUE) #used for centring 

p_stars <- function(p) {
  case_when(
    is.na(p)        ~ "",
    p < 0.001       ~ "☆☆☆",
    p < 0.01        ~ "☆☆",
    p < 0.05        ~ "☆",
    TRUE            ~ ""
  )
}

fmt_p <- function(p) {
  ifelse(is.na(p), NA_character_, format.pval(p, digits = 2, eps = 1e-300))
}

make_bmi_cat <- function(bmi) {
  case_when(
    is.na(bmi) ~ NA_character_, #dont drop them, helps with structure checks later on 
    bmi < 25 ~ "Normal",
    bmi < 30 ~ "Overweight",
    TRUE ~ "Obese"
  )
}

# same as displaying 1st - 99th percentiles just with some fancy name
winsorize <- function(x, probs = c(0.01, 0.99)) {
  qs <- quantile(x, probs = probs, na.rm = TRUE, type = 7)
  pmin(pmax(x, qs[1]), qs[2])
}

## =========================
## PARAMETERS (KEEP ONLY ONCE)
## =========================
lab_vars <- c(
  "Red Blood Cells (Mio/µl)",
  "Basophils abs. (Tsd./µl)",
  "Eosinophils abs. (Tsd./µl)",
  "Immature Granulocytes abs. (Tsd./µl)",
  "White Blood Cells (Tsd./µl)",
  "Lymphocytes abs. (Tsd./µl)",
  "Monocytes abs. (Tsd./µl)",
  "Neutrophils abs. (Tsd./µl)",
  "Platelets (Tsd./µl)",
  "ESR (mm/h)",
  "CRP (mg/l)",
  "RDW (%)",
  "Hemoglobin (g/dl)",
  "Hematocrit (%)",
  "Alkaline Phosphatase (U/l)",
  "Ferritin (µg/l)",
  "Serum Glucose (mg/dl)",
  "Insulin (mIU/l)",
  "GPT/ALT (U/l)",
  "GOT/AST (U/l)",
  "Gamma-GT (U/l)",
  "GFR (ml/min)",
  "Uric Acid (mg/dl)",
  "PDW (fl)",
  "P-LCR (%)",
  "MCH (pg)",
  "MCHC (g/dl)",
  "MCV (fl)",
  "Monocytes (%)",
  "Neutrophils (%)",
  "Basophils (%)",
  "Eosinophils (%)",
  "Lymphocytes (%)",
  "Immature Granulocytes (%)",
  "Total Cholesterol (mg/dl)",
  "Triglycerides (mg/dl)",
  "HDL-C (mg/dl)",
  "Non-HDL Cholesterol (mg/dl)",
  "LDL-C (mg/dl)",
  "LDL/HDL Ratio (ratio)",
  "TG/HDL Ratio"
)

markers_lm <- c(
  "Basophils abs. (Tsd./µl)",
  "Eosinophils abs. (Tsd./µl)",
  "Immature Granulocytes abs. (Tsd./µl)",
  "White Blood Cells (Tsd./µl)",
  "Lymphocytes abs. (Tsd./µl)",
  "Monocytes abs. (Tsd./µl)",
  "Neutrophils abs. (Tsd./µl)",
  "Platelets (Tsd./µl)",
  "ESR (mm/h)",
  "CRP (mg/l)",
  "Ferritin (µg/l)",
  "Serum Glucose (mg/dl)",
  "Insulin (mIU/l)",
  "GFR (ml/min)",
  "Uric Acid (mg/dl)",
  "Total Cholesterol (mg/dl)",
  "Triglycerides (mg/dl)",
  "HDL-C (mg/dl)",
  "Non-HDL Cholesterol (mg/dl)",
  "LDL-C (mg/dl)",
  "LDL/HDL Ratio (ratio)",
  "TG/HDL Ratio",
  "GPT/ALT (U/l)",
  "GOT/AST (U/l)",
  "Gamma-GT (U/l)",
  "Alkaline Phosphatase (U/l)"
)

## =========================
## PARAMETERS (GROUPED)
## =========================
immune_absolute_vars <- c(
  "White Blood Cells (Tsd./µl)†",
  "Neutrophils abs. (Tsd./µl)†",
  "Lymphocytes abs. (Tsd./µl)†",
  "Monocytes abs. (Tsd./µl)†",
  "Eosinophils abs. (Tsd./µl)†",
  "Basophils abs. (Tsd./µl)†",
  "Immature Granulocytes abs. (Tsd./µl)†",
  "Platelets (Tsd./µl)†"
)

immune_percentage_vars <- c(
  "Neutrophils (%)",
  "Lymphocytes (%)",
  "Monocytes (%)",
  "Eosinophils (%)",
  "Basophils (%)",
  "Immature Granulocytes (%)"
)

inflammatory_vars <- c(
  "CRP (mg/l)†",
  "ESR (mm/h)†",
  "Ferritin (µg/l)†"
)

lipid_vars <- c(
  "Total Cholesterol (mg/dl)",
  "LDL-C (mg/dl)",
  "HDL-C (mg/dl)",
  "Non-HDL Cholesterol (mg/dl)",
  "Triglycerides (mg/dl)†"
)

metabolic_vars <- c(
  "Serum Glucose (mg/dl)",
  "Insulin (mIU/l)†",
  "GPT/ALT (U/l)†",
  "GOT/AST (U/l)†",
  "Gamma-GT (U/l)",
  "Alkaline Phosphatase (U/l)",
  "GFR (ml/min)",
  "Uric Acid (mg/dl)"
)

hematology_vars <- c(
  "Red Blood Cells (Mio/µl)",
  "Hemoglobin (g/dl)",
  "Hematocrit (%)",
  "RDW (%)",
  "MCH (pg)",
  "MCHC (g/dl)",
  "MCV (fl)",
  "PDW (fl)",
  "P-LCR (%)"
)

index_vars_needed <- c(
  "Neutrophils abs. (Tsd./µl)",
  "Lymphocytes abs. (Tsd./µl)",
  "Monocytes abs. (Tsd./µl)",
  "Platelets (Tsd./µl)"
)

## =========================
## PARAMETERS ANALYSIS
## =========================
percent_vars <- lab_vars[grepl("%", lab_vars)]

delta_vars <- c(
  "Red Blood Cells (Mio/µl)",
  "Basophils abs. (Tsd./µl)",
  "Eosinophils abs. (Tsd./µl)",
  "Immature Granulocytes abs. (Tsd./µl)",
  "White Blood Cells (Tsd./µl)",
  "Lymphocytes abs. (Tsd./µl)",
  "Monocytes abs. (Tsd./µl)",
  "Neutrophils abs. (Tsd./µl)",
  "Platelets (Tsd./µl)",
  "ESR (mm/h)",
  "CRP (mg/l)",
  "Ferritin (µg/l)",
  "Serum Glucose (mg/dl)",
  "Insulin (mIU/l)",
  "GPT/ALT (U/l)",
  "GOT/AST (U/l)",
  "Gamma-GT (U/l)",
  "GFR (ml/min)",
  "Uric Acid (mg/dl)",
  "Alkaline Phosphatase (U/l)",
  "Total Cholesterol (mg/dl)",
  "Triglycerides (mg/dl)",
  "HDL-C (mg/dl)",
  "Non-HDL Cholesterol (mg/dl)",
  "LDL-C (mg/dl)",
  "TG/HDL Ratio",
  "Hemoglobin (g/dl)",
  "MCH (pg)",
  "MCHC (g/dl)",
  "MCV (fl)",
  "PDW (fl)"
)

## =========================
## READ RAW DATA   
## =========================
FF_laboratory_raw <- read_csv2(   
  "FF_laboratory_ETL2.csv",
  locale = locale(encoding = "ISO-8859-1")
)

FF_fasting <- read_csv("FF_fasting_duration_and_FR_duration.csv") %>%
  mutate(study_id = as.character(study_id))

FF_metadata <- read_excel("C:/Users/megbi/Desktop/BA/Cohort descriptions/FF_metadata_0402026.xlsx") %>%
  select(-sex) %>%
  mutate(study_id = as.character(study_id))

FF_diet <- read_csv("FF_dietary_data_28-11-2025.csv") %>%
  mutate(study_id = as.character(study_id))

## =========================
## CLEAN / STANDARDIZE CORE LABS      
## =========================
FF_laboratory <- FF_laboratory_raw %>%
  select(-lab_id, -case_id, -range_values, -birthdate) %>%
  mutate(
    datetime = as.Date(datetime, format = "%d.%m.%Y"),
    study_id = as.character(study_id),
    variable = ifelse(variable == "High-Sensitivity CRP (mg/l)", "CRP (mg/l)", variable), 
    value = case_when(
      variable == "CRP (mg/l)" & str_detect(value, "<0\\.5") ~ as.character(0.5 / sqrt(2)),
      variable == "CRP (mg/l)" & str_detect(value, "<0\\.16") ~ as.character(0.16 / sqrt(2)),
      variable == "CRP (mg/l)" & str_detect(value, "^<") ~ {
        num_val <- as.numeric(str_extract(value, "[0-9.]+"))
        as.character(num_val / sqrt(2))
      },
      TRUE ~ value
    ),
    value = as.numeric(value)
  ) %>%
  filter(
    grepl("^[0-9]+$", study_id),
    timepoint %in% c("T0","T1","T2"),
    variable %in% lab_vars
  ) %>%
  mutate(timepoint = factor(timepoint, levels = c("T0","T1","T2")))

## =========================
## CLEAN METADATA (ONE ROW PER SUBJECT)
## =========================
FF_metadata_clean <- FF_metadata %>%
  filter(grepl("^[0-9]+$", study_id)) %>%
  group_by(study_id) %>%
  summarise(
    age = first_non_na(age),
    BMI_baseline = first_non_na(BMI_baseline),
    .groups = "drop"
  )

FF_metadata_clean <- FF_metadata_clean %>%
  mutate(
    age = readr::parse_number(as.character(age)),
    BMI_baseline = readr::parse_number(as.character(BMI_baseline))
  )


## =========================
## CLEAN DIET (ONE ROW PER SUBJECT)
## =========================
diet_lookup <- FF_diet %>%
  group_by(study_id) %>%
  summarise(FR_dietary = first_non_na(FR_dietary), .groups = "drop") %>%
  mutate(
    FR_dietary = stringr::str_trim(FR_dietary),
    diet_keto = case_when(
      FR_dietary == "KETO" ~ "Keto",
      FR_dietary %in% c("normal", "Diabete Diet") ~ "Non-Keto",
      TRUE ~ NA_character_
    ),
    diet_keto = factor(diet_keto, levels = c("Non-Keto", "Keto"))
  )

## =========================
## MASTER LONG DATASET (everything joined)
## =========================
FF_laboratory_master <- FF_laboratory %>%
  left_join(FF_fasting, by = "study_id") %>%
  left_join(FF_metadata_clean, by = "study_id") %>%
  left_join(diet_lookup %>% select(study_id, diet_keto), by = "study_id")

FF_laboratory_master <- FF_laboratory_master %>%
  mutate(
    age = readr::parse_number(as.character(age)),
    BMI_baseline = readr::parse_number(as.character(BMI_baseline)),
    fasting_duration_measured = readr::parse_number(as.character(fasting_duration_measured)),
    FR_duration = readr::parse_number(as.character(FR_duration)),
    sex = factor(sex, levels = c("F","M"))
  )

## =========================
## ANCOVA DATASET (LONG, baseline-adjusted; EXCLUDES % and optionally excludes TG/HDL)
## =========================
FF_laboratory_clean <- FF_laboratory %>%
  arrange(study_id, variable, timepoint, datetime) %>%
  group_by(study_id, variable, timepoint) %>%
  summarise(
    value = first(value[!is.na(value)]),
    sex = first(sex),
    .groups = "drop"
  )

baseline_df <- FF_laboratory_clean %>%
  filter(timepoint == "T0") %>%
  transmute(study_id, variable, baseline = value)

FF_ancova <- FF_laboratory_clean %>%
  filter(timepoint %in% c("T1","T2")) %>%
  left_join(baseline_df, by = c("study_id","variable")) %>%
  left_join(FF_metadata_clean, by = "study_id") %>%
  mutate(
    timepoint = factor(timepoint, levels = c("T1","T2")),
    sex = factor(sex, levels = c("F","M"))
  ) %>%
  filter(!variable %in% percent_vars) %>%
  filter(variable != "TG/HDL Ratio")   # remove ratios from ancova if needed 


# =========================
# Clean long lab dataset for CBC / immune cells only
# (earliest sample per subject x marker x timepoint)
# =========================
immune_cells <- c(
  "Neutrophils abs. (Tsd./µl)",
  "Lymphocytes abs. (Tsd./µl)",
  "Monocytes abs. (Tsd./µl)",
  "Eosinophils abs. (Tsd./µl)",
  "Basophils abs. (Tsd./µl)",
  "Immature Granulocytes abs. (Tsd./µl)",
  "White Blood Cells (Tsd./µl)",
  "Platelets (Tsd./µl)"
)

labs_long_cells <- FF_laboratory_master %>%
  filter(timepoint %in% c("T0","T1","T2")) %>%
  filter(variable %in% immune_cells) %>%   # <- keeps it strictly CBC markers
  group_by(study_id, variable, timepoint) %>%
  slice_min(datetime, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  mutate(
    timepoint = factor(timepoint, levels = c("T0","T1","T2")),
    diet_keto = factor(diet_keto, levels = c("Non-Keto","Keto"))
  )

## metabolic health refeeding ANCOVA 
metabolic_index_vars <- c("HOMA_IR", "TyG", "AIP", "non_HDL", "remnant_chol", "remnant_cholesterol")

immune_index_vars <- c(
  "NLR (ratio)", "PLR (ratio)", "LMR (ratio)", "SII (a.u.)",
  "TG/HDL Ratio", "TG/HDL (ratio)", "TG/HDL (Ratio)"
)

# Anything that looks like an index/ratio 
exclude_regex <- "(\\(Ratio\\))|\\(ratio\\))|(^AIP$)|(^TyG$)|(^HOMA)|(^non_HDL$)|(^remnant)|index"

index_vars_exclude <- c(immune_index_vars, metabolic_index_vars)

refeed_base <- FF_laboratory_master %>%
  filter(timepoint %in% c("T1","T2")) %>%
  group_by(study_id, variable, timepoint) %>%
  slice_min(datetime, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  mutate(
    timepoint = factor(timepoint, levels = c("T1","T2")),
    sex       = factor(sex, levels = c("F","M")),
    diet_keto = factor(diet_keto, levels = c("Non-Keto","Keto")),
    FR_duration  = as.numeric(FR_duration),
    age          = as.numeric(age),
    BMI_baseline = as.numeric(BMI_baseline)
  ) %>%
  filter(
    !variable %in% percent_vars,
    !variable %in% index_vars_exclude,
    !grepl(exclude_regex, variable, ignore.case = TRUE)
  ) %>%
  select(study_id, variable, timepoint, value, diet_keto, FR_duration, age, sex, BMI_baseline)


# =========================================================
# Shared helper for all heatmaps
# =========================================================
strip_dagger <- function(x) {
  stringr::str_remove(as.character(x), "†$")
}

marker_order_clean <- c(
  # Lipid
  "Total Cholesterol (mg/dl)",
  "LDL-C (mg/dl)",
  "Non-HDL Cholesterol (mg/dl)",
  "HDL-C (mg/dl)",
  "Triglycerides (mg/dl)",

  # Metabolic / Organ
  "Serum Glucose (mg/dl)",
  "Insulin (mIU/l)",
  "GFR (ml/min)",
  "Uric Acid (mg/dl)",
  "Gamma-GT (U/l)",
  "Alkaline Phosphatase (U/l)",
  "GPT/ALT (U/l)",
  "GOT/AST (U/l)",

  # Inflammation & iron
  "CRP (mg/l)",
  "ESR (mm/h)",
  "Ferritin (µg/l)",

  # Circulating immune cells
  "White Blood Cells (Tsd./µl)",
  "Neutrophils abs. (Tsd./µl)",
  "Lymphocytes abs. (Tsd./µl)",
  "Monocytes abs. (Tsd./µl)",
  "Eosinophils abs. (Tsd./µl)",
  "Basophils abs. (Tsd./µl)",
  "Immature Granulocytes abs. (Tsd./µl)",
  "Platelets (Tsd./µl)"
)

assign_domain_from_variable <- function(variable_vec) {
  variable_clean <- strip_dagger(variable_vec)

  case_when(
    variable_clean %in% c(
      "Total Cholesterol (mg/dl)",
      "LDL-C (mg/dl)",
      "Non-HDL Cholesterol (mg/dl)",
      "HDL-C (mg/dl)",
      "Triglycerides (mg/dl)"
    ) ~ "Lipid",

    variable_clean %in% c(
      "Serum Glucose (mg/dl)",
      "Insulin (mIU/l)",
      "GFR (ml/min)",
      "Uric Acid (mg/dl)",
      "Gamma-GT (U/l)",
      "Alkaline Phosphatase (U/l)",
      "GPT/ALT (U/l)",
      "GOT/AST (U/l)"
    ) ~ "Metabolic / Organ",

    variable_clean %in% c(
      "CRP (mg/l)",
      "ESR (mm/h)",
      "Ferritin (µg/l)"
    ) ~ "Inflammation & iron",

    variable_clean %in% c(
      "White Blood Cells (Tsd./µl)",
      "Neutrophils abs. (Tsd./µl)",
      "Lymphocytes abs. (Tsd./µl)",
      "Monocytes abs. (Tsd./µl)",
      "Eosinophils abs. (Tsd./µl)",
      "Basophils abs. (Tsd./µl)",
      "Immature Granulocytes abs. (Tsd./µl)",
      "Platelets (Tsd./µl)"
    ) ~ "Circulating immune cells",

    TRUE ~ "Other"
  )
}

apply_manual_marker_order <- function(df, variable_col = "Variable") {
  var_sym <- rlang::sym(variable_col)

  marker_levels_present <- df %>%
    distinct(!!var_sym) %>%
    mutate(
      Variable_clean = strip_dagger(as.character(!!var_sym)),
      order_id = match(Variable_clean, marker_order_clean)
    ) %>%
    arrange(is.na(order_id), order_id, !!var_sym) %>%
    pull(!!var_sym)

  df %>%
    mutate(
      marker = factor(!!var_sym, levels = rev(marker_levels_present))
    )
}
```

## 0) Cohort description

```{r cohort}
## =========================================================
## TABLE 1. Cohort description
## Main table only: age, sex, BMI, BMI category,
## fasting duration, refeeding duration, diet group
## =========================================================

# one row per subject
cohort_df <- FF_laboratory_master %>%
  group_by(study_id) %>%
  summarise(
    age = first_non_na(age),
    sex = first_non_na(as.character(sex)),
    BMI_baseline = first_non_na(BMI_baseline),
    fasting_duration_measured = first_non_na(fasting_duration_measured),
    FR_duration = first_non_na(FR_duration),
    diet_keto = first_non_na(as.character(diet_keto)),
    .groups = "drop"
  ) %>%
  mutate(
    sex = factor(sex, levels = c("F", "M")),
    diet_keto = factor(diet_keto, levels = c("Non-Keto", "Keto")),
    BMI_cat = factor(
      make_bmi_cat(BMI_baseline),
      levels = c("Normal", "Overweight", "Obese")
    )
  )

# helpers
fmt_num <- function(x, digits = 1) {
  ifelse(is.na(x), NA_character_, format(round(x, digits), nsmall = digits))
}

fmt_mean_sd <- function(x, digits = 1) {
  x <- x[!is.na(x)]
  if (length(x) == 0) return(NA_character_)
  paste0(fmt_num(mean(x), digits), " ± ", fmt_num(sd(x), digits))
}

fmt_median_iqr <- function(x, digits = 1) {
  x <- x[!is.na(x)]
  if (length(x) == 0) return(NA_character_)
  q <- quantile(x, probs = c(0.25, 0.5, 0.75), na.rm = TRUE)
  paste0(
    fmt_num(q[2], digits),
    " [", fmt_num(q[1], digits), ", ", fmt_num(q[3], digits), "]"
  )
}

fmt_n_pct <- function(n, denom, digits = 1) {
  if (is.na(denom) || denom == 0) return(NA_character_)
  paste0(n, " (", fmt_num(100 * n / denom, digits), "%)")
}

# denominators
# basically to make sure you dont account for missing subjects
n_total <- nrow(cohort_df)
n_diet_nonmiss <- sum(!is.na(cohort_df$diet_keto))
n_bmi_nonmiss <- sum(!is.na(cohort_df$BMI_cat))

# final table
table1_cohort <- tibble(
  Characteristic = c(
    "Number of subjects",
    "Age (years), mean ± SD",
    "Age (years), median [IQR]",
    "Baseline BMI (kg/m²), mean ± SD",
    "Baseline BMI (kg/m²), median [IQR]",
    "BMI category: Normal, n (%)",
    "BMI category: Overweight, n (%)",
    "BMI category: Obese, n (%)",
    "Fasting duration (days), mean ± SD",
    "Fasting duration (days), median [IQR]",
    "Refeeding duration (days), mean ± SD",
    "Refeeding duration (days), median [IQR]",
    "Female, n (%)",
    "Male, n (%)",
    "Diet: Non-Keto, n (%)",
    "Diet: Keto, n (%)"
  ),
  `Total cohort` = c(
    n_total,
    fmt_mean_sd(cohort_df$age, 1),
    fmt_median_iqr(cohort_df$age, 1),
    fmt_mean_sd(cohort_df$BMI_baseline, 1),
    fmt_median_iqr(cohort_df$BMI_baseline, 1),
    fmt_n_pct(sum(cohort_df$BMI_cat == "Normal", na.rm = TRUE), n_bmi_nonmiss),
    fmt_n_pct(sum(cohort_df$BMI_cat == "Overweight", na.rm = TRUE), n_bmi_nonmiss),
    fmt_n_pct(sum(cohort_df$BMI_cat == "Obese", na.rm = TRUE), n_bmi_nonmiss),
    fmt_mean_sd(cohort_df$fasting_duration_measured, 1),
    fmt_median_iqr(cohort_df$fasting_duration_measured, 1),
    fmt_mean_sd(cohort_df$FR_duration, 1),
    fmt_median_iqr(cohort_df$FR_duration, 1),
    fmt_n_pct(sum(cohort_df$sex == "F", na.rm = TRUE), n_total),
    fmt_n_pct(sum(cohort_df$sex == "M", na.rm = TRUE), n_total),
    fmt_n_pct(sum(cohort_df$diet_keto == "Non-Keto", na.rm = TRUE), n_diet_nonmiss),
    fmt_n_pct(sum(cohort_df$diet_keto == "Keto", na.rm = TRUE), n_diet_nonmiss)
  )
)

table1_cohort %>%
  knitr::kable(
    caption = "Table 1. Cohort characteristics.",
    col.names = c("Characteristic", "Total cohort"),
    align = c("l", "l")
  ) %>%
  kableExtra::kable_styling(
    bootstrap_options = c("striped", "hover"),
    full_width = FALSE
  )

# optional save
dir.create("ModelTables", showWarnings = FALSE)
readr::write_csv(table1_cohort, "ModelTables/Table1_cohort_description.csv")
```

## 1) Descriptive Trajectories: Response to Overall Intervention

```{r summary-boxplots (totpop,BMI)}
## =========================
## summary boxplots + median lines for all stratification (distribution) as well as whole pop
## 
## @vivi just to see the changes induced by the intervention, what rebounds, decrease or increases
## significance will be tested later of course
## now we just look at the data at face value and compare with basic biology
## =========================
markers_to_plot <- c(
  immune_absolute_vars,
  immune_percentage_vars,
  inflammatory_vars,
  lipid_vars,
  metabolic_vars,
  hematology_vars
) %>%
  strip_dagger() %>%
  unique()

markers_to_plot <- intersect(markers_to_plot, unique(FF_laboratory_master$variable))

log1p_markers <- c(
  immune_absolute_vars,
  inflammatory_vars
) %>%
  strip_dagger() %>%
  unique()

plot_marker_box_medians <- function(
  data,
  marker,
  use_log1p = FALSE,
  cap_quantiles = c(0.01, 0.99)
) {
  df <- data %>%
    filter(variable == marker, timepoint %in% c("T0", "T1", "T2"), !is.na(value)) %>%
    mutate(
      timepoint = factor(timepoint, levels = c("T0", "T1", "T2")),
      BMI_cat = factor(
        make_bmi_cat(BMI_baseline),
        levels = c("Normal", "Overweight", "Obese")
      )
    )

  if (nrow(df) == 0) return(NULL)

  # cap per marker
  lo <- quantile(df$value, cap_quantiles[1], na.rm = TRUE)
  hi <- quantile(df$value, cap_quantiles[2], na.rm = TRUE)

  df <- df %>%
    mutate(
      value_cap = pmin(pmax(value, lo), hi),
      value_plot = if (use_log1p) log1p(value_cap) else value_cap
    )

  # medians on plotted scale
  med_all <- df %>%
    group_by(timepoint) %>%
    summarise(med = median(value_plot, na.rm = TRUE), .groups = "drop") %>%
    mutate(group = "All")

  med_bmi <- df %>%
    filter(!is.na(BMI_cat)) %>%
    group_by(timepoint, BMI_cat) %>%
    summarise(med = median(value_plot, na.rm = TRUE), .groups = "drop")

  # means on capped original scale, then transformed for plotting
  mean_all <- df %>%
    group_by(timepoint) %>%
    summarise(mean_cap = mean(value_cap, na.rm = TRUE), .groups = "drop") %>%
    mutate(mean_plot = if (use_log1p) log1p(mean_cap) else mean_cap)

  # n labels
  n_tp <- df %>%
    group_by(timepoint) %>%
    summarise(n = n(), .groups = "drop") %>%
    mutate(lbl = paste0("n=", n))

  n_bmi <- df %>%
    filter(!is.na(BMI_cat)) %>%
    group_by(timepoint, BMI_cat) %>%
    summarise(n = n(), .groups = "drop") %>%
    mutate(lbl = paste0("n=", n))

  y_top <- max(df$value_plot, na.rm = TRUE)
  y_range <- diff(range(df$value_plot, na.rm = TRUE))
  if (!is.finite(y_range) || y_range == 0) y_range <- 1

  n_tp <- n_tp %>%
  mutate(y = y_top + 0.04 * y_range)  # was 0.10 or 0.15

  bmi_levels <- levels(df$BMI_cat)
  n_bmi <- n_bmi %>%
    mutate(
      bmi_i = match(BMI_cat, bmi_levels),
      y = y_top - (bmi_i - 1) * 0.06 * y_range
    )

  ggplot(df, aes(x = timepoint, y = value_plot)) +
    geom_boxplot(
      outlier.shape = NA,
      width = 0.6,
      alpha = 0.8
    ) +
    geom_line(
      data = med_all,
      aes(x = timepoint, y = med, group = group),
      linewidth = 1.2,
      alpha = 1.0,
      inherit.aes = FALSE
    ) +
    geom_point(
      data = med_all,
      aes(x = timepoint, y = med),
      size = 2.4,
      alpha = 1.0,
      inherit.aes = FALSE
    ) +
    geom_point(
      data = mean_all,
      aes(x = timepoint, y = mean_plot),
      shape = 18,
      size = 2.8,
      color = "red",
      inherit.aes = FALSE
    ) +
    geom_line(
      data = med_bmi,
      aes(x = timepoint, y = med, group = BMI_cat, linetype = BMI_cat),
      linewidth = 0.9,
      alpha = 0.45,
      inherit.aes = FALSE
    ) +
    geom_point(
      data = med_bmi,
      aes(x = timepoint, y = med, group = BMI_cat),
      size = 1.6,
      alpha = 0.45,
      inherit.aes = FALSE
    ) +
    geom_text(
  data = n_tp,
  aes(x = timepoint, y = y, label = lbl),
  inherit.aes = FALSE,
  size = 3,
  fontface = "bold",
  hjust = -0.2,
  nudge_x = 0.1
) +
geom_text(
  data = n_bmi,
  aes(x = timepoint, y = y, label = paste0(BMI_cat, ": ", lbl)),
  inherit.aes = FALSE,
  size = 2.5,
  alpha = 0.75,
  hjust = -0.1,
  nudge_x = 0.1
) +
    labs(
      title = marker,
      x = "Timepoint",
      y = if (use_log1p) "Value (log1p; Winsorised 1st–99th pct)" else "Value (Winsorised 1st–99th pct)",
      linetype = "BMI category"
    ) +
    theme_bw() +
    theme(
      legend.position = "bottom",
      panel.grid.minor = element_blank(),
      plot.margin = margin(8, 8, 8, 8)
    )
}

dir.create("MarkerPlots_Medians", showWarnings = FALSE)

plots_out <- vector("list", length(markers_to_plot))
names(plots_out) <- markers_to_plot

for (m in markers_to_plot) {
  p <- plot_marker_box_medians(
    FF_laboratory_master,
    marker = m,
    use_log1p = m %in% log1p_markers
  )

  plots_out[[m]] <- p

  if (!is.null(p)) {
    safe_name <- gsub("[^A-Za-z0-9]+", "_", m)
    ggsave(
      filename = file.path("MarkerPlots_Medians", paste0("Marker_", safe_name, ".png")),
      plot = p,
      width = 7,
      height = 5,
      dpi = 300
    )
  }
}
```

## 2) Fasting Intervention

```{r fasting-effects}
## =====================================================================
## FASTING ANALYSES (T0 vs T1) - CLEAN TEMPLATE
## - same marker universe + ratio exclusions
## - log-transform selected markers BEFORE modeling (applies to T0 & T1)
## - centers continuous covariates within each marker sample (ALL models)
## - prints LME (timepoint shift), LME (timepoint x fasting duration),
##   ANCOVA (T1 ~ T0 + covariates), ANCOVA + BMI interaction, ANCOVA + Fast interaction
## - FDR computed within each term across markers
## =====================================================================
min_n <- 20

## -------------------------
## 0) Marker lists + log list
## -------------------------
markers_fasting <- markers_lm %>%
  setdiff(c("TG/HDL Ratio", "LDL/HDL Ratio (ratio)"))

markers_log <- c(
  "CRP (mg/l)",
  "Ferritin (µg/l)",
  "Triglycerides (mg/dl)",
  "Insulin (mIU/l)",
  "ESR (mm/h)",
  "GPT/ALT (U/l)",
  "GOT/AST (U/l)",
  "Neutrophils abs. (Tsd./µl)",
  "Lymphocytes abs. (Tsd./µl)",
  "Monocytes abs. (Tsd./µl)",
  "Eosinophils abs. (Tsd./µl)",
  "Basophils abs. (Tsd./µl)",
  "Immature Granulocytes abs. (Tsd./µl)",
  "White Blood Cells (Tsd./µl)",
  "Platelets (Tsd./µl)"
)

safe_log <- function(x) ifelse(!is.na(x) & x > 0, log(x), NA_real_)

center_in_sample <- function(x) x - mean(x, na.rm = TRUE)

fmt_ci <- function(lo, hi, digits = 3) {
  ifelse(is.na(lo) | is.na(hi), NA_character_,
         paste0("[", round(lo, digits), ", ", round(hi, digits), "]"))
}

## -------------------------
## 1) Term sets (expected)
## -------------------------
terms_lme_time <- c("(Intercept)", "timepointT1", "age_c", "sexM", "BMI_c")

terms_lme_fast <- c(
  "(Intercept)",
  "timepointT1",
  "fast_c",
  "timepointT1:fast_c",
  "age_c",
  "sexM",
  "BMI_c"
)

terms_ancova_main <- c("T0_c", "fast_c", "BMI_c", "age_c", "sexM")

terms_ancova_BMIint <- c("T0_c", "BMI_c", "T0_c:BMI_c", "fast_c", "age_c", "sexM")

terms_ancova_FASTint <- c("T0_c", "fast_c", "T0_c:fast_c", "BMI_c", "age_c", "sexM")

## -------------------------
## 2) Pretty term names + ordering
## -------------------------
term_pretty_fasting <- function(term) {
  dplyr::case_when(
    term == "(Intercept)"      ~ "Intercept",
    term == "timepointT1"      ~ "Timepoint: T1 vs T0",
    term == "fast_c"           ~ "Fasting duration (days), centered",
    term == "timepointT1:fast_c" ~ "Timepoint × fasting duration (Δ per day)",
    term == "T0_c"             ~ "Baseline (T0), centered",
    term == "BMI_c"            ~ "Baseline BMI (kg/m²), centered",
    term == "age_c"            ~ "Age (years), centered",
    term == "sexM"             ~ "Sex: Male (vs Female)",
    term == "T0_c:BMI_c"       ~ "T0 × BMI (baseline dependency differs by BMI)",
    term == "T0_c:fast_c"      ~ "T0 × fasting duration (baseline dependency differs by duration)",
    TRUE ~ term
  )
}

term_order_fasting <- function(term) {
  dplyr::case_when(
    term %in% c("timepointT1:fast_c") ~ 1L,
    term %in% c("timepointT1")        ~ 2L,
    term %in% c("T0_c:BMI_c", "T0_c:fast_c") ~ 3L,
    term %in% c("T0_c")               ~ 4L,
    term %in% c("fast_c")             ~ 5L,
    term %in% c("BMI_c")              ~ 6L,
    term %in% c("age_c")              ~ 7L,
    term %in% c("sexM")               ~ 8L,
    term == "(Intercept)"             ~ 9L,
    TRUE ~ 99L
  )
}

## -------------------------
## 3) Build ONE clean long dataset (T0/T1)
## -------------------------
fasting_long <- FF_laboratory_master %>%
  filter(
    variable %in% markers_fasting,
    timepoint %in% c("T0","T1"),
    !is.na(value)
  ) %>%
  group_by(study_id, variable, timepoint) %>%
  slice_min(datetime, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  mutate(
    timepoint = factor(timepoint, levels = c("T0","T1")),
    sex       = factor(sex, levels = c("F","M")),

    age  = readr::parse_number(as.character(age)),
    BMI_baseline = readr::parse_number(as.character(BMI_baseline)),
    fasting_duration_measured = readr::parse_number(as.character(fasting_duration_measured)),

    value = ifelse(variable %in% markers_log, safe_log(value), value),

    Variable = ifelse(variable %in% markers_log, paste0(variable, "†"), variable)
  ) %>%
  filter(!is.na(value))

## =====================================================================
## 4.1) Fasting average shift (LME) - ALL TERMS
## Model: value ~ timepoint + age_c + sex + BMI_c + (1|study_id)
## =====================================================================

lme_time_tbl <- fasting_long %>%
  group_by(variable) %>%
  group_modify(~{
    df <- .x %>%
      drop_na(value, timepoint, age, sex, BMI_baseline)

    n_subj <- n_distinct(df$study_id)
    if (n_subj < min_n) {
      return(tibble(
        term = terms_lme_time,
        estimate = NA_real_, std.error = NA_real_,
        conf.low = NA_real_, conf.high = NA_real_,
        p.value = NA_real_,
        n_subjects = n_subj
      ))
    }

    df <- df %>%
      mutate(
        age_c = center_in_sample(age),
        BMI_c = center_in_sample(BMI_baseline)
      )

    fit <- lmer(value ~ timepoint + age_c + sex + BMI_c + (1|study_id), data = df)

    broom.mixed::tidy(fit, effects = "fixed", conf.int = TRUE) %>%
      filter(term %in% terms_lme_time) %>%
      mutate(n_subjects = n_subj)
  }) %>%
  ungroup() %>%
  mutate(p_raw = p.value) %>%
  group_by(term) %>%
  mutate(p_fdr = p.adjust(p_raw, method = "BH")) %>%
  ungroup()

lme_time_table <- lme_time_tbl %>%
  left_join(fasting_long %>% distinct(variable, Variable), by = "variable") %>%
  mutate(
    Term = term_pretty_fasting(term),
    Term_order = term_order_fasting(term),
    Estimate = round(estimate, 3),
    SE = round(std.error, 3),
    `95% CI` = fmt_ci(conf.low, conf.high, digits = 3),
    `P (raw)` = fmt_p(p_raw),
    `P (FDR)` = fmt_p(p_fdr),
    `Signif.` = p_stars(p_fdr)
  ) %>%
  select(
    Variable, Term, Estimate, SE, `95% CI`,
    `P (raw)`, `P (FDR)`, `Signif.`,
    `n (subjects)` = n_subjects,
    Term_order, p_fdr
  ) %>%
  arrange(Term_order, is.na(p_fdr), p_fdr, Variable) %>%
  select(-Term_order, -p_fdr)

lme_time_table %>%
  kbl(
    caption = "Fasting LME (all fixed effects): value ~ timepoint + centered age + sex + centered BMI + (1|study_id). FDR computed within each term across markers.",
    align = "l"
  ) %>%
  kable_styling(bootstrap_options = c("striped","hover"), full_width = FALSE) %>%
  footnote(general = "† Log-transformed prior to centering and modeling.") %>%
  scroll_box(width = "100%", height = "650px")

dir.create("ModelTables", showWarnings = FALSE)
write_csv(lme_time_table, file.path("ModelTables", "LME_fasting_all_terms_T0T1.csv"))

## =====================================================================
## 4.2) Fasting duration modulation (LME) - ALL TERMS
## Model: value ~ timepoint * fast_c + age_c + sex + BMI_c + (1|study_id)
## =====================================================================

lme_fastmod_tbl <- fasting_long %>%
  group_by(variable) %>%
  group_modify(~{
    df <- .x %>%
      drop_na(value, timepoint, fasting_duration_measured, age, sex, BMI_baseline)

    n_subj <- n_distinct(df$study_id)
    if (n_subj < min_n) {
      return(tibble(
        term = terms_lme_fast,
        estimate = NA_real_, std.error = NA_real_,
        conf.low = NA_real_, conf.high = NA_real_,
        p.value = NA_real_,
        n_subjects = n_subj
      ))
    }

    df <- df %>%
      mutate(
        fast_c = center_in_sample(fasting_duration_measured),
        age_c  = center_in_sample(age),
        BMI_c  = center_in_sample(BMI_baseline)
      )

    fit <- lmer(value ~ timepoint * fast_c + age_c + sex + BMI_c + (1|study_id), data = df)

    broom.mixed::tidy(fit, effects = "fixed", conf.int = TRUE) %>%
      filter(term %in% terms_lme_fast) %>%
      mutate(n_subjects = n_subj)
  }) %>%
  ungroup() %>%
  mutate(p_raw = p.value) %>%
  group_by(term) %>%
  mutate(p_fdr = p.adjust(p_raw, method = "BH")) %>%
  ungroup()

lme_fastmod_table <- lme_fastmod_tbl %>%
  left_join(fasting_long %>% distinct(variable, Variable), by = "variable") %>%
  mutate(
    Term = term_pretty_fasting(term),
    Term_order = term_order_fasting(term),
    Estimate = round(estimate, 3),
    SE = round(std.error, 3),
    `95% CI` = fmt_ci(conf.low, conf.high, digits = 3),
    `P (raw)` = fmt_p(p_raw),
    `P (FDR)` = fmt_p(p_fdr),
    `Signif.` = p_stars(p_fdr)
  ) %>%
  select(
    Variable, Term, Estimate, SE, `95% CI`,
    `P (raw)`, `P (FDR)`, `Signif.`,
    `n (subjects)` = n_subjects,
    Term_order, p_fdr
  ) %>%
  arrange(Term_order, is.na(p_fdr), p_fdr, Variable) %>%
  select(-Term_order, -p_fdr)

lme_fastmod_table %>%
  kbl(
    caption = "Fasting LME (all fixed effects): value ~ timepoint × centered fasting duration + centered age + sex + centered BMI + (1|study_id). FDR computed within each term across markers.",
    align = "l"
  ) %>%
  kable_styling(bootstrap_options = c("striped","hover"), full_width = FALSE) %>%
  footnote(general = "† Log-transformed prior to centering and modeling.") %>%
  scroll_box(width = "100%", height = "650px")

write_csv(lme_fastmod_table, file.path("ModelTables", "LME_fasting_timepointXfast_all_terms_T0T1.csv"))

## =====================================================================
## 4.3) Fasting ANCOVA (main effects)
## Model: T1 ~ T0_c + fast_c + BMI_c + age_c + sex
## (centering is within marker sample; log already applied in fasting_long)
## =====================================================================

fasting_wide <- fasting_long %>%
  select(study_id, variable, Variable, timepoint, value, fasting_duration_measured, age, sex, BMI_baseline) %>%
  pivot_wider(names_from = timepoint, values_from = value) %>%
  filter(!is.na(T0), !is.na(T1)) %>%
  group_by(variable) %>%
  mutate(
    T0_c   = center_in_sample(T0),
    fast_c = center_in_sample(fasting_duration_measured),
    age_c  = center_in_sample(age),
    BMI_c  = center_in_sample(BMI_baseline)
  ) %>%
  ungroup()

ancova_main_tbl <- fasting_wide %>%
  group_by(variable) %>%
  group_modify(~{
    df <- .x %>% drop_na(T1, T0_c, fast_c, BMI_c, age_c, sex)

    n_subj <- n_distinct(df$study_id)
    if (n_subj < min_n) {
      return(tibble(
        term = terms_ancova_main,
        estimate = NA_real_, std.error = NA_real_,
        conf.low = NA_real_, conf.high = NA_real_,
        p.value = NA_real_,
        n_subjects = n_subj
      ))
    }

    fit <- lm(T1 ~ T0_c + fast_c + BMI_c + age_c + sex, data = df)

    broom::tidy(fit, conf.int = TRUE) %>%
      filter(term %in% terms_ancova_main) %>%
      mutate(n_subjects = n_subj)
  }) %>%
  ungroup() %>%
  mutate(p_raw = p.value) %>%
  group_by(term) %>%
  mutate(p_fdr = p.adjust(p_raw, method = "BH")) %>%
  ungroup()

ancova_main_table <- ancova_main_tbl %>%
  left_join(fasting_wide %>% distinct(variable, Variable), by = "variable") %>%
  mutate(
    Term = term_pretty_fasting(term),
    Term_order = term_order_fasting(term),
    Estimate = round(estimate, 3),
    SE = round(std.error, 3),
    `95% CI` = fmt_ci(conf.low, conf.high, digits = 3),
    `P (raw)` = fmt_p(p_raw),
    `P (FDR)` = fmt_p(p_fdr),
    `Signif.` = p_stars(p_fdr)
  ) %>%
  select(
    Variable, Term, Estimate, SE, `95% CI`,
    `P (raw)`, `P (FDR)`, `Signif.`,
    `n (subjects)` = n_subjects,
    Term_order, p_fdr
  ) %>%
  arrange(Term_order, is.na(p_fdr), p_fdr, Variable) %>%
  select(-Term_order, -p_fdr)

ancova_main_table %>%
  kbl(
    caption = "Fasting ANCOVA (per marker): T1 adjusted for centered T0 + centered fasting duration + centered BMI + centered age + sex. FDR computed within each term across markers.",
    align = "l"
  ) %>%
  kable_styling(bootstrap_options = c("striped","hover"), full_width = FALSE) %>%
  footnote(general = "† Log-transformed prior to centering and modeling.") %>%
  scroll_box(width = "100%", height = "650px")

write_csv(ancova_main_table, file.path("ModelTables", "LM_fasting_ANCOVA_main_all_terms_T1onT0.csv"))

## =====================================================================
## 4.4) Fasting ANCOVA + BMI interaction
## Model: T1 ~ T0_c * BMI_c + fast_c + age_c + sex
## =====================================================================

ancova_BMIint_tbl <- fasting_wide %>%
  group_by(variable) %>%
  group_modify(~{
    df <- .x %>% drop_na(T1, T0_c, BMI_c, fast_c, age_c, sex)

    n_subj <- n_distinct(df$study_id)
    if (n_subj < min_n) {
      return(tibble(
        term = terms_ancova_BMIint,
        estimate = NA_real_, std.error = NA_real_,
        conf.low = NA_real_, conf.high = NA_real_,
        p.value = NA_real_,
        n_subjects = n_subj
      ))
    }

    fit <- lm(T1 ~ T0_c * BMI_c + fast_c + age_c + sex, data = df)

    broom::tidy(fit, conf.int = TRUE) %>%
      filter(term %in% terms_ancova_BMIint) %>%
      mutate(n_subjects = n_subj)
  }) %>%
  ungroup() %>%
  mutate(p_raw = p.value) %>%
  group_by(term) %>%
  mutate(p_fdr = p.adjust(p_raw, method = "BH")) %>%
  ungroup()

ancova_BMIint_table <- ancova_BMIint_tbl %>%
  left_join(fasting_wide %>% distinct(variable, Variable), by = "variable") %>%
  mutate(
    Term = term_pretty_fasting(term),
    Term_order = term_order_fasting(term),
    Estimate = round(estimate, 3),
    SE = round(std.error, 3),
    `95% CI` = fmt_ci(conf.low, conf.high, digits = 3),
    `P (raw)` = fmt_p(p_raw),
    `P (FDR)` = fmt_p(p_fdr),
    `Signif.` = p_stars(p_fdr)
  ) %>%
  select(
    Variable, Term, Estimate, SE, `95% CI`,
    `P (raw)`, `P (FDR)`, `Signif.`,
    `n (subjects)` = n_subjects,
    Term_order, p_fdr
  ) %>%
  arrange(Term_order, is.na(p_fdr), p_fdr, Variable) %>%
  select(-Term_order, -p_fdr)

ancova_BMIint_table %>%
  kbl(
    caption = "Fasting ANCOVA + BMI interaction (per marker): T1 ~ centered T0 × centered BMI + centered fasting duration + centered age + sex. FDR within term across markers.",
    align = "l"
  ) %>%
  kable_styling(bootstrap_options = c("striped","hover"), full_width = FALSE) %>%
  footnote(general = "† Log-transformed prior to centering and modeling.") %>%
  scroll_box(width = "100%", height = "650px")

write_csv(ancova_BMIint_table, file.path("ModelTables", "LM_fasting_ANCOVA_T0xBMI_all_terms_T1onT0.csv"))

## =====================================================================
## 4.5) Fasting ANCOVA + fasting duration interaction
## Model: T1 ~ T0_c * fast_c + BMI_c + age_c + sex
## =====================================================================

ancova_FASTint_tbl <- fasting_wide %>%
  group_by(variable) %>%
  group_modify(~{
    df <- .x %>% drop_na(T1, T0_c, fast_c, BMI_c, age_c, sex)

    n_subj <- n_distinct(df$study_id)
    if (n_subj < min_n) {
      return(tibble(
        term = terms_ancova_FASTint,
        estimate = NA_real_, std.error = NA_real_,
        conf.low = NA_real_, conf.high = NA_real_,
        p.value = NA_real_,
        n_subjects = n_subj
      ))
    }

    fit <- lm(T1 ~ T0_c * fast_c + BMI_c + age_c + sex, data = df)

    broom::tidy(fit, conf.int = TRUE) %>%
      filter(term %in% terms_ancova_FASTint) %>%
      mutate(n_subjects = n_subj)
  }) %>%
  ungroup() %>%
  mutate(p_raw = p.value) %>%
  group_by(term) %>%
  mutate(p_fdr = p.adjust(p_raw, method = "BH")) %>%
  ungroup()

ancova_FASTint_table <- ancova_FASTint_tbl %>%
  left_join(fasting_wide %>% distinct(variable, Variable), by = "variable") %>%
  mutate(
    Term = term_pretty_fasting(term),
    Term_order = term_order_fasting(term),
    Estimate = round(estimate, 3),
    SE = round(std.error, 3),
    `95% CI` = fmt_ci(conf.low, conf.high, digits = 3),
    `P (raw)` = fmt_p(p_raw),
    `P (FDR)` = fmt_p(p_fdr),
    `Signif.` = p_stars(p_fdr)
  ) %>%
  select(
    Variable, Term, Estimate, SE, `95% CI`,
    `P (raw)`, `P (FDR)`, `Signif.`,
    `n (subjects)` = n_subjects,
    Term_order, p_fdr
  ) %>%
  arrange(Term_order, is.na(p_fdr), p_fdr, Variable) %>%
  select(-Term_order, -p_fdr)

ancova_FASTint_table %>%
  kbl(
    caption = "Fasting ANCOVA + fasting duration interaction (per marker): T1 ~ centered T0 × centered fasting duration + centered BMI + centered age + sex. FDR within term across markers.",
    align = "l"
  ) %>%
  kable_styling(bootstrap_options = c("striped","hover"), full_width = FALSE) %>%
  footnote(general = "† Log-transformed prior to centering and modeling.") %>%
  scroll_box(width = "100%", height = "650px")

write_csv(ancova_FASTint_table, file.path("ModelTables", "LM_fasting_ANCOVA_T0xFast_all_terms_T1onT0.csv"))
```



## 3) Refeeding Intervention

```{r sats-refeed, message=FALSE, warning=FALSE}
## =====================================================================
## REFEEDING ANALYSES (T1 vs T2) - FASTING-STYLE TEMPLATE (CLEAN)
## - same marker universe and ratio exclusions as fasting
## - log-transform selected markers BEFORE modeling (applies to T1 & T2)
## - centers continuous covariates within each marker sample
## - shows ALL fixed-effect terms (LMEs)
## - FDR computed within each term across markers
## =====================================================================

min_n <- 20

## -------------------------
## 0) Consistent marker lists
## -------------------------
# Same universe and ratio exclusions as fasting
markers_refeed <- markers_lm %>%
  setdiff(c("TG/HDL Ratio", "LDL/HDL Ratio (ratio)"))

# Same log group as fasting (you already validated positivity in audit)
markers_log <- c(
  "CRP (mg/l)",
  "Ferritin (µg/l)",
  "Triglycerides (mg/dl)",
  "Insulin (mIU/l)",
  "ESR (mm/h)",
  "GPT/ALT (U/l)",
  "GOT/AST (U/l)",
  "Neutrophils abs. (Tsd./µl)",
  "Lymphocytes abs. (Tsd./µl)",
  "Monocytes abs. (Tsd./µl)",
  "Eosinophils abs. (Tsd./µl)",
  "Basophils abs. (Tsd./µl)",
  "Immature Granulocytes abs. (Tsd./µl)",
  "White Blood Cells (Tsd./µl)",
  "Platelets (Tsd./µl)"
)

safe_log <- function(x) if_else(!is.na(x) & x > 0, log(x), NA_real_)

## -------------------------
## 1) Term sets (expected)
## -------------------------
terms_lme_time <- c("(Intercept)", "timepointT2", "age_c", "sexM", "BMI_c")

terms_lme_fr <- c(
  "(Intercept)",
  "timepointT2",
  "FR_c",
  "timepointT2:FR_c",
  "age_c",
  "sexM",
  "BMI_c"
)

terms_ancova <- c("T1_c", "diet_ketoKeto", "FR_c", "age_c", "sexM", "BMI_c")

## -------------------------
## 2) Pretty terms + ordering
## -------------------------
term_pretty_refeed <- function(term) {
  dplyr::case_when(
    term == "(Intercept)"        ~ "Intercept",
    term == "timepointT2"        ~ "Timepoint: T2 vs T1",
    term == "FR_c"               ~ "Refeeding duration (days), centered",
    term == "timepointT2:FR_c"   ~ "Timepoint × refeeding duration (Δ per day)",
    term == "T1_c"               ~ "End of fasting (T1), centered",
    term == "diet_ketoKeto"      ~ "Diet: Keto (vs Non-Keto)",
    term == "age_c"              ~ "Age (years), centered",
    term == "sexM"               ~ "Sex: Male (vs Female)",
    term == "BMI_c"              ~ "Baseline BMI (kg/m²), centered",
    TRUE ~ term
  )
}

term_order_refeed <- function(term) {
  dplyr::case_when(
    term == "timepointT2:FR_c" ~ 1L,
    term == "timepointT2"      ~ 2L,
    term == "T1_c"             ~ 3L,
    term == "diet_ketoKeto"    ~ 4L,
    term == "FR_c"             ~ 5L,
    term == "BMI_c"            ~ 6L,
    term == "age_c"            ~ 7L,
    term == "sexM"             ~ 8L,
    term == "(Intercept)"      ~ 9L,
    TRUE ~ 99L
  )
}

## -------------------------
## 3) Build ONE clean long dataset (T1/T2)
## -------------------------
refeed_long <- refeed_base %>%
  filter(
    variable %in% markers_refeed,
    timepoint %in% c("T1","T2"),
    !is.na(value)
  ) %>%
  mutate(
    timepoint = factor(timepoint, levels = c("T1","T2")),
    sex       = factor(sex, levels = c("F","M")),
    diet_keto = factor(diet_keto, levels = c("Non-Keto","Keto")),  # ensures diet_ketoKeto exists

    age          = as.numeric(age),
    BMI_baseline = as.numeric(BMI_baseline),
    FR_duration  = as.numeric(FR_duration),

    # log BEFORE modeling (applies at both timepoints)
    value = if_else(variable %in% markers_log, safe_log(value), value),

    # dagger label for tables
    Variable = if_else(variable %in% markers_log, paste0(variable, "†"), variable)
  ) %>%
  filter(!is.na(value))  # drop NA created by safe_log (should be none per your audit)

## =====================================================================
## 8.1) Refeeding average shift (LME) - ALL TERMS
## Model: value ~ timepoint + age_c + sex + BMI_c + (1|study_id)
## =====================================================================

lme_time_tbl <- refeed_long %>%
  group_by(variable) %>%
  group_modify(~{
    df <- .x %>% drop_na(value, timepoint, age, sex, BMI_baseline)

    n_subj <- n_distinct(df$study_id)
    if (n_subj < min_n) {
      return(tibble(
        term = terms_lme_time,
        estimate = NA_real_, std.error = NA_real_,
        conf.low = NA_real_, conf.high = NA_real_,
        p.value = NA_real_,
        n_subjects = n_subj
      ))
    }

    # center within marker sample
    df <- df %>%
      mutate(
        age_c = age - mean(age, na.rm = TRUE),
        BMI_c = BMI_baseline - mean(BMI_baseline, na.rm = TRUE)
      )

    fit <- lmer(value ~ timepoint + age_c + sex + BMI_c + (1 | study_id), data = df)

    broom.mixed::tidy(fit, effects = "fixed", conf.int = TRUE) %>%
      filter(term %in% terms_lme_time) %>%
      mutate(n_subjects = n_subj)
  }) %>%
  ungroup() %>%
  mutate(p_raw = p.value) %>%
  group_by(term) %>%
  mutate(p_fdr = p.adjust(p_raw, method = "BH")) %>%
  ungroup()

lme_time_table <- lme_time_tbl %>%
  left_join(refeed_long %>% distinct(variable, Variable), by = "variable") %>%
  mutate(
    Term = term_pretty_refeed(term),
    Term_order = term_order_refeed(term),
    Estimate = round(estimate, 3),
    SE = round(std.error, 3),
    `95% CI` = ifelse(
      is.na(conf.low) | is.na(conf.high),
      NA_character_,
      paste0("[", round(conf.low, 3), ", ", round(conf.high, 3), "]")
    ),
    `P (raw)` = fmt_p(p_raw),
    `P (FDR)` = fmt_p(p_fdr),
    `Signif.` = p_stars(p_fdr)
  ) %>%
  select(
    Variable, Term, Estimate, SE, `95% CI`,
    `P (raw)`, `P (FDR)`, `Signif.`,
    `n (subjects)` = n_subjects,
    Term_order, p_fdr
  ) %>%
  arrange(Term_order, is.na(p_fdr), p_fdr, Variable) %>%
  select(-Term_order, -p_fdr)

lme_time_table %>%
  kbl(
    caption = "Refeeding LME (all fixed effects): value ~ timepoint + age + sex + BMI + (1|study_id). FDR computed within each term across markers.",
    align = "l"
  ) %>%
  kable_styling(bootstrap_options = c("striped","hover"), full_width = FALSE) %>%
  footnote(general = "† Log-transformed prior to centering and modeling.") %>%
  scroll_box(width = "100%", height = "650px")

dir.create("ModelTables", showWarnings = FALSE)
write_csv(lme_time_table, file.path("ModelTables", "LME_refeed_all_terms_T1T2.csv"))

## =====================================================================
## 8.2) Refeeding duration modulation (LME) - ALL TERMS
## Model: value ~ timepoint * FR_c + age_c + sex + BMI_c + (1|study_id)
## =====================================================================

lme_frmod_tbl <- refeed_long %>%
  group_by(variable) %>%
  group_modify(~{
    df <- .x %>% drop_na(value, timepoint, FR_duration, age, sex, BMI_baseline)

    n_subj <- n_distinct(df$study_id)
    if (n_subj < min_n) {
      return(tibble(
        term = terms_lme_fr,
        estimate = NA_real_, std.error = NA_real_,
        conf.low = NA_real_, conf.high = NA_real_,
        p.value = NA_real_,
        n_subjects = n_subj
      ))
    }

    df <- df %>%
      mutate(
        FR_c  = FR_duration  - mean(FR_duration, na.rm = TRUE),
        age_c = age          - mean(age, na.rm = TRUE),
        BMI_c = BMI_baseline - mean(BMI_baseline, na.rm = TRUE)
      )

    fit <- lmer(value ~ timepoint * FR_c + age_c + sex + BMI_c + (1 | study_id), data = df)

    broom.mixed::tidy(fit, effects = "fixed", conf.int = TRUE) %>%
      filter(term %in% terms_lme_fr) %>%
      mutate(n_subjects = n_subj)
  }) %>%
  ungroup() %>%
  mutate(p_raw = p.value) %>%
  group_by(term) %>%
  mutate(p_fdr = p.adjust(p_raw, method = "BH")) %>%
  ungroup()

lme_frmod_table <- lme_frmod_tbl %>%
  left_join(refeed_long %>% distinct(variable, Variable), by = "variable") %>%
  mutate(
    Term = term_pretty_refeed(term),
    Term_order = term_order_refeed(term),
    Estimate = round(estimate, 3),
    SE = round(std.error, 3),
    `95% CI` = ifelse(
      is.na(conf.low) | is.na(conf.high),
      NA_character_,
      paste0("[", round(conf.low, 3), ", ", round(conf.high, 3), "]")
    ),
    `P (raw)` = fmt_p(p_raw),
    `P (FDR)` = fmt_p(p_fdr),
    `Signif.` = p_stars(p_fdr)
  ) %>%
  select(
    Variable, Term, Estimate, SE, `95% CI`,
    `P (raw)`, `P (FDR)`, `Signif.`,
    `n (subjects)` = n_subjects,
    Term_order, p_fdr
  ) %>%
  arrange(Term_order, is.na(p_fdr), p_fdr, Variable) %>%
  select(-Term_order, -p_fdr)

lme_frmod_table %>%
  kbl(
    caption = "Refeeding LME (all fixed effects): value ~ timepoint * centered FR_duration + age + sex + BMI + (1|study_id). FDR computed within each term across markers.",
    align = "l"
  ) %>%
  kable_styling(bootstrap_options = c("striped","hover"), full_width = FALSE) %>%
  footnote(general = "† Log-transformed prior to centering and modeling.") %>%
  scroll_box(width = "100%", height = "650px")

write_csv(lme_frmod_table, file.path("ModelTables", "LME_refeed_timepointXFR_all_terms_T1T2.csv"))

## =====================================================================
## 8.3) Refeeding ANCOVA (fasting-style)
## Model: T2 ~ T1_c + diet_keto + FR_c + age_c + sex + BMI_c
## - log-transform is applied to BOTH T1 and T2 via refeed_long BEFORE widening
## - centers within marker sample
## - shows ALL terms in term_keep; FDR within term across markers
## =====================================================================

refeed_wide <- refeed_long %>%
  select(study_id, variable, Variable, timepoint, value, diet_keto, FR_duration, age, sex, BMI_baseline) %>%
  pivot_wider(names_from = timepoint, values_from = value) %>%
  filter(!is.na(T1), !is.na(T2)) %>%
  group_by(variable) %>%
  mutate(
    T1_c  = T1 - mean(T1, na.rm = TRUE),
    FR_c  = FR_duration - mean(FR_duration, na.rm = TRUE),
    age_c = age - mean(age, na.rm = TRUE),
    BMI_c = BMI_baseline - mean(BMI_baseline, na.rm = TRUE)
  ) %>%
  ungroup()

ancova_tbl_refeed <- refeed_wide %>%
  group_by(variable) %>%
  group_modify(~{
    df <- .x %>% drop_na(T1_c, T2, diet_keto, FR_c, age_c, sex, BMI_c)

    n_subj <- n_distinct(df$study_id)
    if (n_subj < min_n) {
      return(tibble(
        term = terms_ancova,
        estimate = NA_real_, std.error = NA_real_,
        conf.low = NA_real_, conf.high = NA_real_,
        p.value = NA_real_,
        n_subjects = n_subj
      ))
    }

    fit <- lm(T2 ~ T1_c + diet_keto + FR_c + age_c + sex + BMI_c, data = df)

    broom::tidy(fit, conf.int = TRUE) %>%
      filter(term %in% terms_ancova) %>%
      mutate(n_subjects = n_subj)
  }) %>%
  ungroup() %>%
  mutate(p_raw = p.value) %>%
  group_by(term) %>%
  mutate(p_fdr = p.adjust(p_raw, method = "BH")) %>%
  ungroup()

ancova_table_refeed <- ancova_tbl_refeed %>%
  left_join(refeed_wide %>% distinct(variable, Variable), by = "variable") %>%
  mutate(
    Term = term_pretty_refeed(term),
    Term_order = term_order_refeed(term),
    Estimate = round(estimate, 3),
    SE = round(std.error, 3),
    `95% CI` = ifelse(
      is.na(conf.low) | is.na(conf.high),
      NA_character_,
      paste0("[", round(conf.low, 3), ", ", round(conf.high, 3), "]")
    ),
    `P (raw)` = fmt_p(p_raw),
    `P (FDR)` = fmt_p(p_fdr),
    `Signif.` = p_stars(p_fdr)
  ) %>%
  select(
    Variable, Term, Estimate, SE, `95% CI`,
    `P (raw)`, `P (FDR)`, `Signif.`,
    `n (subjects)` = n_subjects,
    Term_order, p_fdr
  ) %>%
  arrange(Term_order, is.na(p_fdr), p_fdr, Variable) %>%
  select(-Term_order, -p_fdr)

ancova_table_refeed %>%
  kbl(
    caption = "Refeeding ANCOVA (per marker): T2 adjusted for centered T1 + diet + centered FR_duration (+ age, sex, BMI). FDR computed within each term across markers.",
    align = "l"
  ) %>%
  kable_styling(bootstrap_options = c("striped","hover"), full_width = FALSE) %>%
  footnote(general = "† Log-transformed prior to centering and modeling.") %>%
  scroll_box(width = "100%", height = "650px")

write_csv(ancova_table_refeed, file.path("ModelTables", "LM_refeed_ANCOVA_centered_all_terms_T2onT1.csv"))

## =====================================================================
## 8.4) Refeeding ANCOVA + Keto interaction
## Model: T2 ~ T1_c * diet_keto + FR_c + age_c + sex + BMI_c
## Question: Does the T1→T2 relationship differ by Keto vs Non-Keto?
## (i.e., baseline dependency differs by diet group)
## =====================================================================

terms_ancova_keto_int <- c(
  "T1_c",
  "diet_ketoKeto",
  "T1_c:diet_ketoKeto",
  "FR_c",
  "age_c",
  "sexM",
  "BMI_c"
)

term_pretty_refeed2 <- function(term) {
  dplyr::case_when(
    term == "T1_c"                 ~ "End of fasting (T1), centered",
    term == "diet_ketoKeto"        ~ "Diet: Keto (vs Non-Keto)",
    term == "T1_c:BMI_c"         ~ "T1 × BMI (baseline dependency differs by BMI)",
    term == "T1_c:diet_ketoKeto"   ~ "T1 × Diet (baseline dependency differs by diet)",
    term == "FR_c"                 ~ "Refeeding duration (days), centered",
    term == "age_c"                ~ "Age (years), centered",
    term == "sexM"                 ~ "Sex: Male (vs Female)",
    term == "BMI_c"                ~ "Baseline BMI (kg/m²), centered",
    term == "(Intercept)"          ~ "Intercept",
    TRUE ~ term
  )
}

term_order_refeed2 <- function(term) {
  dplyr::case_when(
    term == "T1_c:diet_ketoKeto" ~ 1L,
    term == "T1_c:BMI_c"         ~ 1L,   # used in next model too
    term == "T1_c"               ~ 2L,
    term == "diet_ketoKeto"      ~ 3L,
    term == "BMI_c"              ~ 4L,
    term == "FR_c"               ~ 5L,
    term == "age_c"              ~ 6L,
    term == "sexM"               ~ 7L,
    term == "(Intercept)"        ~ 8L,
    TRUE ~ 99L
  )
}

ancova_tbl_refeed_keto_int <- refeed_wide %>%
  group_by(variable) %>%
  group_modify(~{
    df <- .x %>% drop_na(T1_c, T2, diet_keto, FR_c, age_c, sex, BMI_c)

    n_subj <- n_distinct(df$study_id)
    if (n_subj < min_n) {
      return(tibble(
        term = terms_ancova_keto_int,
        estimate = NA_real_, std.error = NA_real_,
        conf.low = NA_real_, conf.high = NA_real_,
        p.value = NA_real_,
        n_subjects = n_subj
      ))
    }

    fit <- lm(T2 ~ T1_c * diet_keto + FR_c + age_c + sex + BMI_c, data = df)

    broom::tidy(fit, conf.int = TRUE) %>%
      filter(term %in% terms_ancova_keto_int) %>%
      mutate(n_subjects = n_subj)
  }) %>%
  ungroup() %>%
  mutate(p_raw = p.value) %>%
  group_by(term) %>%
  mutate(p_fdr = p.adjust(p_raw, method = "BH")) %>%
  ungroup()

ancova_table_refeed_keto_int <- ancova_tbl_refeed_keto_int %>%
  left_join(refeed_wide %>% distinct(variable, Variable), by = "variable") %>%
  mutate(
    Term = term_pretty_refeed2(term),
    Term_order = term_order_refeed2(term),
    Estimate = round(estimate, 3),
    SE = round(std.error, 3),
    `95% CI` = ifelse(
      is.na(conf.low) | is.na(conf.high),
      NA_character_,
      paste0("[", round(conf.low, 3), ", ", round(conf.high, 3), "]")
    ),
    `P (raw)` = fmt_p(p_raw),
    `P (FDR)` = fmt_p(p_fdr),
    `Signif.` = p_stars(p_fdr)
  ) %>%
  select(
    Variable, Term, Estimate, SE, `95% CI`,
    `P (raw)`, `P (FDR)`, `Signif.`,
    `n (subjects)` = n_subjects,
    Term_order, p_fdr
  ) %>%
  arrange(Term_order, is.na(p_fdr), p_fdr, Variable) %>%
  select(-Term_order, -p_fdr)

ancova_table_refeed_keto_int %>%
  kbl(
    caption = "Refeeding ANCOVA + diet interaction: T2 ~ T1_c * diet_keto + FR_c + age_c + sex + BMI_c. FDR computed within each term across markers.",
    align = "l"
  ) %>%
  kable_styling(bootstrap_options = c("striped","hover"), full_width = FALSE) %>%
  footnote(general = "† Log-transformed prior to centering and modeling.") %>%
  scroll_box(width = "100%", height = "650px")

write_csv(
  ancova_table_refeed_keto_int,
  file.path("ModelTables", "LM_refeed_ANCOVA_T2onT1_dietInteraction.csv")
)


## =====================================================================
## 8.5) Refeeding ANCOVA + BMI interaction
## Model: T2 ~ T1_c * BMI_c + diet_keto + FR_c + age_c + sex
## Question: Does baseline dependency (T1→T2) differ by BMI?
## =====================================================================

terms_ancova_bmi_int <- c(
  "T1_c",
  "BMI_c",
  "T1_c:BMI_c",
  "diet_ketoKeto",
  "FR_c",
  "age_c",
  "sexM"
)

ancova_tbl_refeed_bmi_int <- refeed_wide %>%
  group_by(variable) %>%
  group_modify(~{
    df <- .x %>% drop_na(T1_c, T2, diet_keto, FR_c, age_c, sex, BMI_c)

    n_subj <- n_distinct(df$study_id)
    if (n_subj < min_n) {
      return(tibble(
        term = terms_ancova_bmi_int,
        estimate = NA_real_, std.error = NA_real_,
        conf.low = NA_real_, conf.high = NA_real_,
        p.value = NA_real_,
        n_subjects = n_subj
      ))
    }

    fit <- lm(T2 ~ T1_c * BMI_c + diet_keto + FR_c + age_c + sex, data = df)

    broom::tidy(fit, conf.int = TRUE) %>%
      filter(term %in% terms_ancova_bmi_int) %>%
      mutate(n_subjects = n_subj)
  }) %>%
  ungroup() %>%
  mutate(p_raw = p.value) %>%
  group_by(term) %>%
  mutate(p_fdr = p.adjust(p_raw, method = "BH")) %>%
  ungroup()

ancova_table_refeed_bmi_int <- ancova_tbl_refeed_bmi_int %>%
  left_join(refeed_wide %>% distinct(variable, Variable), by = "variable") %>%
  mutate(
    Term = term_pretty_refeed2(term),
    Term_order = term_order_refeed2(term),
    Estimate = round(estimate, 3),
    SE = round(std.error, 3),
    `95% CI` = ifelse(
      is.na(conf.low) | is.na(conf.high),
      NA_character_,
      paste0("[", round(conf.low, 3), ", ", round(conf.high, 3), "]")
    ),
    `P (raw)` = fmt_p(p_raw),
    `P (FDR)` = fmt_p(p_fdr),
    `Signif.` = p_stars(p_fdr)
  ) %>%
  select(
    Variable, Term, Estimate, SE, `95% CI`,
    `P (raw)`, `P (FDR)`, `Signif.`,
    `n (subjects)` = n_subjects,
    Term_order, p_fdr
  ) %>%
  arrange(Term_order, is.na(p_fdr), p_fdr, Variable) %>%
  select(-Term_order, -p_fdr)

ancova_table_refeed_bmi_int %>%
  kbl(
    caption = "Refeeding ANCOVA + BMI interaction: T2 ~ T1_c * BMI_c + diet_keto + FR_c + age_c + sex. FDR computed within each term across markers.",
    align = "l"
  ) %>%
  kable_styling(bootstrap_options = c("striped","hover"), full_width = FALSE) %>%
  footnote(general = "† Log-transformed prior to centering and modeling.") %>%
  scroll_box(width = "100%", height = "650px")

write_csv(
  ancova_table_refeed_bmi_int,
  file.path("ModelTables", "LM_refeed_ANCOVA_T2onT1_BMIInteraction.csv")
)

## =====================================================================
## 8.6) Refeeding ANCOVA + FR interaction (baseline dependency differs by refeeding duration)
## Model (per marker): T2 ~ T1_c * FR_c + diet_keto + BMI_c + age_c + sex
## - uses refeed_long (already log-transformed for markers_log) BEFORE widening
## - centers within marker sample
## - shows ALL terms; FDR computed within each term across markers
## =====================================================================

# --- terms expected in this model
terms_ancova_FRint <- c(
  "T1_c",
  "FR_c",
  "T1_c:FR_c",
  "diet_ketoKeto",
  "BMI_c",
  "age_c",
  "sexM"
)

# --- pretty names + ordering (add FR interaction)
term_pretty_ancova_FRint <- function(term) {
  dplyr::case_when(
    term == "T1_c"          ~ "End of fasting (T1), centered",
    term == "FR_c"          ~ "Refeeding duration (days), centered",
    term == "T1_c:FR_c"     ~ "T1 × refeeding duration (baseline dependency mod.)",
    term == "diet_ketoKeto" ~ "Diet: Keto (vs Non-Keto)",
    term == "BMI_c"         ~ "Baseline BMI (kg/m²), centered",
    term == "age_c"         ~ "Age (years), centered",
    term == "sexM"          ~ "Sex: Male (vs Female)",
    term == "(Intercept)"   ~ "Intercept",
    TRUE ~ term
  )
}

term_order_ancova_FRint <- function(term) {
  dplyr::case_when(
    term == "T1_c:FR_c"     ~ 1L,
    term == "T1_c"          ~ 2L,
    term == "FR_c"          ~ 3L,
    term == "diet_ketoKeto" ~ 4L,
    term == "BMI_c"         ~ 5L,
    term == "age_c"         ~ 6L,
    term == "sexM"          ~ 7L,
    term == "(Intercept)"   ~ 8L,
    TRUE ~ 99L
  )
}

# --- wide dataset from refeed_long (log already applied there)
refeed_wide_FRint <- refeed_long %>%
  select(study_id, variable, Variable, timepoint, value,
         diet_keto, FR_duration, age, sex, BMI_baseline) %>%
  pivot_wider(names_from = timepoint, values_from = value) %>%
  filter(!is.na(T1), !is.na(T2)) %>%
  group_by(variable) %>%
  mutate(
    T1_c  = T1 - mean(T1, na.rm = TRUE),
    FR_c  = FR_duration - mean(FR_duration, na.rm = TRUE),
    age_c = age - mean(age, na.rm = TRUE),
    BMI_c = BMI_baseline - mean(BMI_baseline, na.rm = TRUE)
  ) %>%
  ungroup()

ancova_tbl_refeed_FRint <- refeed_wide_FRint %>%
  group_by(variable) %>%
  group_modify(~{
    df <- .x %>% drop_na(T2, T1_c, FR_c, diet_keto, BMI_c, age_c, sex)

    n_subj <- n_distinct(df$study_id)
    if (n_subj < min_n) {
      return(tibble(
        term = c("(Intercept)", terms_ancova_FRint),
        estimate = NA_real_, std.error = NA_real_,
        conf.low = NA_real_, conf.high = NA_real_,
        p.value = NA_real_,
        n_subjects = n_subj
      ))
    }

    fit <- lm(T2 ~ T1_c * FR_c + diet_keto + BMI_c + age_c + sex, data = df)

    broom::tidy(fit, conf.int = TRUE) %>%
      filter(term %in% c("(Intercept)", terms_ancova_FRint)) %>%
      mutate(n_subjects = n_subj)
  }) %>%
  ungroup() %>%
  mutate(p_raw = p.value) %>%
  group_by(term) %>%
  mutate(p_fdr = p.adjust(p_raw, method = "BH")) %>%
  ungroup()

ancova_table_refeed_FRint <- ancova_tbl_refeed_FRint %>%
  left_join(refeed_wide_FRint %>% distinct(variable, Variable), by = "variable") %>%
  mutate(
    Term = term_pretty_ancova_FRint(term),
    Term_order = term_order_ancova_FRint(term),

    Estimate = round(estimate, 3),
    SE = round(std.error, 3),
    `95% CI` = ifelse(
      is.na(conf.low) | is.na(conf.high),
      NA_character_,
      paste0("[", round(conf.low, 3), ", ", round(conf.high, 3), "]")
    ),
    `P (raw)` = fmt_p(p_raw),
    `P (FDR)` = fmt_p(p_fdr),
    `Signif.` = p_stars(p_fdr)
  ) %>%
  select(
    Variable, Term, Estimate, SE, `95% CI`,
    `P (raw)`, `P (FDR)`, `Signif.`,
    `n (subjects)` = n_subjects,
    Term_order, p_fdr
  ) %>%
  arrange(Term_order, is.na(p_fdr), p_fdr, Variable) %>%
  select(-Term_order, -p_fdr)

ancova_table_refeed_FRint %>%
  kbl(
    caption = "Refeeding ANCOVA + FR interaction (per marker): T2 ~ T1_c * FR_c + diet + BMI + age + sex. FDR computed within each term across markers.",
    align = "l"
  ) %>%
  kable_styling(bootstrap_options = c("striped","hover"), full_width = FALSE) %>%
  footnote(general = "† Log-transformed prior to centering and modeling.") %>%
  scroll_box(width = "100%", height = "650px")

write_csv(
  ancova_table_refeed_FRint,
  file.path("ModelTables", "LM_refeed_ANCOVA_T2onT1_T1xFR_centered_all_terms.csv")
)
```

## 4) Heterogeneity and bidirectionality of circulating immune cells

```{r immune-waterfall, message=FALSE, warning=FALSE}
suppressPackageStartupMessages({
  library(dplyr)
  library(tidyr)
  library(ggplot2)
  library(forcats)
  library(stringr)
})

## ============================================================
## Build wide immune-cell datasets for waterfall plots
## ============================================================

# FASTING: T0 -> T1
cells_fast_wide <- labs_long_cells %>%
  filter(timepoint %in% c("T0", "T1")) %>%
  select(study_id, variable, timepoint, value) %>%
  pivot_wider(
    names_from = timepoint,
    values_from = value
  ) %>%
  filter(!is.na(T0), !is.na(T1)) %>%
  mutate(delta_T1_T0 = T1 - T0)

# REFEEDING: T1 -> T2
cells_refeed_wide <- labs_long_cells %>%
  filter(timepoint %in% c("T1", "T2")) %>%
  select(study_id, variable, timepoint, value) %>%
  pivot_wider(
    names_from = timepoint,
    values_from = value
  ) %>%
  filter(!is.na(T1), !is.na(T2)) %>%
  mutate(delta_T2_T1 = T2 - T1)

## -------------------------
## Pretty facet labels
## -------------------------
clean_cell_label <- function(x) {
  x %>%
    str_replace_all(" abs\\. \\(Tsd\\./µl\\)", "") %>%
    str_replace_all(" \\(Tsd\\./µl\\)", "") %>%
    str_replace_all("([a-z])([A-Z])", "\\1 \\2") %>%
    str_to_title()
}

marker_order_cells <- c(
  "White Blood Cells (Tsd./µl)",
  "Neutrophils abs. (Tsd./µl)",
  "Lymphocytes abs. (Tsd./µl)",
  "Monocytes abs. (Tsd./µl)",
  "Eosinophils abs. (Tsd./µl)",
  "Basophils abs. (Tsd./µl)",
  "Immature Granulocytes abs. (Tsd./µl)",
  "Platelets (Tsd./µl)"
)

## -------------------------
## Helper: waterfall plot
## -------------------------
plot_waterfall_delta <- function(df_wide, delta_name, y_label) {

  df_long <- df_wide %>%
    select(study_id, variable, all_of(delta_name)) %>%
    filter(!is.na(.data[[delta_name]])) %>%
    mutate(
      variable = factor(variable, levels = marker_order_cells),
      variable_label = clean_cell_label(as.character(variable))
    ) %>%
    group_by(variable) %>%
    arrange(.data[[delta_name]], .by_group = TRUE) %>%
    mutate(rank_within = row_number()) %>%
    ungroup()

  ggplot(df_long, aes(x = rank_within, y = .data[[delta_name]])) +
    geom_hline(yintercept = 0, linetype = "dashed", linewidth = 0.8) +
    geom_col(width = 0.9, alpha = 0.85, fill = "grey45") +
    facet_wrap(~ variable_label, scales = "free_y", ncol = 3) +
    labs(
      title = NULL,
      x = "Individuals (ordered by change)",
      y = y_label
    ) +
    scale_y_continuous(expand = expansion(mult = c(0.05, 0.08))) +
    theme_bw(base_size = 12) +
    theme(
      panel.grid.minor = element_blank(),
      panel.grid.major.x = element_blank(),
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank(),
      axis.title.x = element_text(
        face = "plain",
        margin = margin(t = 12)
      ),
      axis.title.y = element_text(
        face = "plain",
        margin = margin(r = 10)
      ),
      strip.background = element_rect(fill = "grey92", color = "black"),
      strip.text = element_text(face = "bold", size = 11),
      plot.title = element_blank()
    )
}

## -------------------------
## Make plots
## -------------------------
p_water_fast <- plot_waterfall_delta(
  cells_fast_wide,
  delta_name = "delta_T1_T0",
  y_label = expression(Delta~"Cell Count (Tsd./"*mu*"l, T1 - T0)")
)

p_water_refeed <- plot_waterfall_delta(
  cells_refeed_wide,
  delta_name = "delta_T2_T1",
  y_label = expression(Delta~"Cell Count (Tsd./"*mu*"l, T2 - T1)")
)

print(p_water_fast)
print(p_water_refeed)

dir.create("ResultsFigures", showWarnings = FALSE)

ggsave(
  "ResultsFigures/waterfall_immune_cells_fasting.png",
  p_water_fast,
  width = 10.5,
  height = 8,
  dpi = 600
)

ggsave(
  "ResultsFigures/waterfall_immune_cells_fasting.pdf",
  p_water_fast,
  width = 10.5,
  height = 8
)

ggsave(
  "ResultsFigures/waterfall_immune_cells_refeeding.png",
  p_water_refeed,
  width = 10.5,
  height = 8,
  dpi = 600
)

ggsave(
  "ResultsFigures/waterfall_immune_cells_refeeding.pdf",
  p_water_refeed,
  width = 10.5,
  height = 8
)
```

## 5) Average Phase Shift

```{r lme-heatmap-fasting-refeed-average, message=FALSE, warning=FALSE}
suppressPackageStartupMessages({
  library(tidyverse)
  library(readr)
  library(scales)
  library(stringr)
})

# ---- paths
base_dir <- "C:/Users/megbi/Desktop/BA/Differential Blood Count Graphs/Master script 2.0/ModelTables"
fast_path   <- file.path(base_dir, "LME_fasting_all_terms_T0T1.csv")
refeed_path <- file.path(base_dir, "LME_refeed_all_terms_T1T2.csv")

fast_tab   <- read_csv(fast_path, show_col_types = FALSE)
refeed_tab <- read_csv(refeed_path, show_col_types = FALSE)

# ---- helper: remove dagger for matching only
strip_dagger <- function(x) {
  stringr::str_remove(as.character(x), "†$")
}

# ---- MANUAL marker order (same structure/order you want to reuse later)
marker_order_clean <- c(
  # Lipid
  "Total Cholesterol (mg/dl)",
  "LDL-C (mg/dl)",
  "Non-HDL Cholesterol (mg/dl)",
  "HDL-C (mg/dl)",
  "Triglycerides (mg/dl)",

  # Metabolic / Organ
  "Serum Glucose (mg/dl)",
  "Insulin (mIU/l)",
  "GFR (ml/min)",
  "Uric Acid (mg/dl)",
  "Gamma-GT (U/l)",
  "Alkaline Phosphatase (U/l)",
  "GPT/ALT (U/l)",
  "GOT/AST (U/l)",

  # Inflammation & iron
  "CRP (mg/l)",
  "ESR (mm/h)",
  "Ferritin (µg/l)",

  # Circulating immune cells
  "White Blood Cells (Tsd./µl)",
  "Neutrophils abs. (Tsd./µl)",
  "Lymphocytes abs. (Tsd./µl)",
  "Monocytes abs. (Tsd./µl)",
  "Eosinophils abs. (Tsd./µl)",
  "Basophils abs. (Tsd./µl)",
  "Immature Granulocytes abs. (Tsd./µl)",
  "Platelets (Tsd./µl)"
)

# ---- robust extractor
extract_shift <- function(df, phase_label, term_regex) {
  out <- df %>%
    mutate(Term = as.character(Term)) %>%
    filter(grepl(term_regex, Term)) %>%
    mutate(
      phase = phase_label,
      Estimate = as.numeric(Estimate),
      signif_star = case_when(
        is.na(`Signif.`) ~ "",
        str_trim(as.character(`Signif.`)) == "ns" ~ "",
        TRUE ~ as.character(`Signif.`)
      )
    )

  if (nrow(out) == 0) {
    stop(paste0("No rows matched term_regex='", term_regex, "' for phase='", phase_label, "'."))
  }

  out
}

fast_shift <- extract_shift(fast_tab,   "Fasting",   "T1\\s*vs\\s*T0")
refe_shift <- extract_shift(refeed_tab, "Refeeding", "T2\\s*vs\\s*T1")

lme_shift_all <- bind_rows(fast_shift, refe_shift) %>%
  mutate(
    phase = factor(phase, levels = c("Fasting", "Refeeding")),
    contrast = if_else(phase == "Fasting", "T1 vs T0", "T2 vs T1"),
    contrast = factor(contrast, levels = c("T1 vs T0", "T2 vs T1")),

    Variable_clean = strip_dagger(Variable)
  )

# ---- domain classification using clean labels
domain_order <- c("Lipid", "Metabolic / Organ", "Inflammation & iron", "Circulating immune cells", "Other")

lme_shift_all <- lme_shift_all %>%
  mutate(
    domain = case_when(
      Variable_clean %in% c(
        "Total Cholesterol (mg/dl)",
        "LDL-C (mg/dl)",
        "Non-HDL Cholesterol (mg/dl)",
        "HDL-C (mg/dl)",
        "Triglycerides (mg/dl)"
      ) ~ "Lipid",

      Variable_clean %in% c(
        "Serum Glucose (mg/dl)",
        "Insulin (mIU/l)",
        "GFR (ml/min)",
        "Uric Acid (mg/dl)",
        "Gamma-GT (U/l)",
        "Alkaline Phosphatase (U/l)",
        "GPT/ALT (U/l)",
        "GOT/AST (U/l)"
      ) ~ "Metabolic / Organ",

      Variable_clean %in% c(
        "CRP (mg/l)",
        "ESR (mm/h)",
        "Ferritin (µg/l)"
      ) ~ "Inflammation & iron",

      Variable_clean %in% c(
        "White Blood Cells (Tsd./µl)",
        "Neutrophils abs. (Tsd./µl)",
        "Lymphocytes abs. (Tsd./µl)",
        "Monocytes abs. (Tsd./µl)",
        "Eosinophils abs. (Tsd./µl)",
        "Basophils abs. (Tsd./µl)",
        "Immature Granulocytes abs. (Tsd./µl)",
        "Platelets (Tsd./µl)"
      ) ~ "Circulating immune cells",

      TRUE ~ "Other"
    ),
    domain = factor(domain, levels = domain_order)
  )

# ---- manual marker ordering
marker_levels_present <- lme_shift_all %>%
  distinct(Variable, Variable_clean) %>%
  mutate(order_id = match(Variable_clean, marker_order_clean)) %>%
  arrange(is.na(order_id), order_id, Variable) %>%
  pull(Variable)

lme_shift_all <- lme_shift_all %>%
  mutate(
    marker = factor(Variable, levels = rev(marker_levels_present))
  )

# ---- scaling
lims <- range(lme_shift_all$Estimate, na.rm = TRUE)

signed_sqrt_trans <- scales::trans_new(
  name = "signed_sqrt",
  transform = function(x) sign(x) * sqrt(abs(x)),
  inverse   = function(x) sign(x) * (abs(x)^2)
)

# ---- plot
p_heat2 <- ggplot(lme_shift_all, aes(x = contrast, y = marker, fill = Estimate)) +
  geom_tile(color = "grey92", linewidth = 0.25) +
  geom_text(aes(label = signif_star), size = 4.7) +
  facet_grid(
    domain ~ phase,
    scales = "free",
    space  = "free_y",
    switch = "y"
  ) +
  scale_y_discrete(expand = expansion(mult = 0.04)) +
  scale_x_discrete(drop = TRUE, expand = expansion(mult = 0.04)) +
  scale_fill_gradient2(
    low = "#0f766e",
    mid = "#ddd5c6",
    high = "#f59e0b",
    midpoint = 0,
    limits = lims,
    oob = scales::squish,
    trans = signed_sqrt_trans,
    breaks = c(-23.0, 0, 4.8),
    name = "Estimate"
  ) +
  labs(x = NULL, y = NULL) +
  theme_bw() +
  theme(
    strip.placement = "outside",
    strip.background = element_blank(),
    strip.text.y.left = element_text(angle = 0, face = "bold", size = 11),
    strip.text.x = element_text(face = "bold", size = 12),
    panel.spacing.y = unit(0.5, "lines"),
    panel.spacing.x = unit(0.5, "lines"),
    panel.grid = element_blank(),
    plot.margin = margin(5, 5, 5, 5),
    axis.text.y = element_text(size = 10),
    axis.text.x = element_text(size = 11),
    legend.position = "right"
  )

p_heat2

# ---- save
out_dir <- "C:/Users/megbi/Desktop/BA/Differential Blood Count Graphs/Master script 2.0/ResultsFigures"
dir.create(out_dir, recursive = TRUE, showWarnings = FALSE)

ggsave(file.path(out_dir, "heatmap_LME_fasting_vs_refeed_two_columns_fixedOrder.png"),
       p_heat2, width = 10.5, height = 10.5, dpi = 300)

ggsave(file.path(out_dir, "heatmap_LME_fasting_vs_refeed_two_columns_fixedOrder.pdf"),
       p_heat2, width = 10.5, height = 10.5)

summary(lme_shift_all$Estimate)
```

## 6) Duration dependence

```{r lme-heatmap-fasting-refeed-dose, message=FALSE, warning=FALSE}
suppressPackageStartupMessages({
  library(tidyverse)
  library(readr)
  library(stringr)
  library(scales)
})

# ---- paths
base_dir <- "C:/Users/megbi/Desktop/BA/Differential Blood Count Graphs/Master script 2.0/ModelTables"
fast_int_path   <- file.path(base_dir, "LME_fasting_timepointXfast_all_terms_T0T1.csv")
refeed_int_path <- file.path(base_dir, "LME_refeed_timepointXFR_all_terms_T1T2.csv")

fast_int_tab   <- read_csv(fast_int_path, show_col_types = FALSE)
refeed_int_tab <- read_csv(refeed_int_path, show_col_types = FALSE)

# ---- helper: remove dagger for matching only
strip_dagger <- function(x) {
  stringr::str_remove(as.character(x), "†$")
}

# ---- manual marker order (same as section 8)
marker_order_clean <- c(
  # Lipid
  "Total Cholesterol (mg/dl)",
  "LDL-C (mg/dl)",
  "Non-HDL Cholesterol (mg/dl)",
  "HDL-C (mg/dl)",
  "Triglycerides (mg/dl)",

  # Metabolic / Organ
  "Serum Glucose (mg/dl)",
  "Insulin (mIU/l)",
  "GFR (ml/min)",
  "Uric Acid (mg/dl)",
  "Gamma-GT (U/l)",
  "Alkaline Phosphatase (U/l)",
  "GPT/ALT (U/l)",
  "GOT/AST (U/l)",

  # Inflammation & iron
  "CRP (mg/l)",
  "ESR (mm/h)",
  "Ferritin (µg/l)",

  # Circulating immune cells
  "White Blood Cells (Tsd./µl)",
  "Neutrophils abs. (Tsd./µl)",
  "Lymphocytes abs. (Tsd./µl)",
  "Monocytes abs. (Tsd./µl)",
  "Eosinophils abs. (Tsd./µl)",
  "Basophils abs. (Tsd./µl)",
  "Immature Granulocytes abs. (Tsd./µl)",
  "Platelets (Tsd./µl)"
)

# ---- helper: domain assignment
assign_domain_from_variable <- function(x) {
  x_clean <- strip_dagger(x)

  case_when(
    x_clean %in% c(
      "Total Cholesterol (mg/dl)",
      "LDL-C (mg/dl)",
      "Non-HDL Cholesterol (mg/dl)",
      "HDL-C (mg/dl)",
      "Triglycerides (mg/dl)"
    ) ~ "Lipid",

    x_clean %in% c(
      "Serum Glucose (mg/dl)",
      "Insulin (mIU/l)",
      "GFR (ml/min)",
      "Uric Acid (mg/dl)",
      "Gamma-GT (U/l)",
      "Alkaline Phosphatase (U/l)",
      "GPT/ALT (U/l)",
      "GOT/AST (U/l)"
    ) ~ "Metabolic / Organ",

    x_clean %in% c(
      "CRP (mg/l)",
      "ESR (mm/h)",
      "Ferritin (µg/l)"
    ) ~ "Inflammation & iron",

    x_clean %in% c(
      "White Blood Cells (Tsd./µl)",
      "Neutrophils abs. (Tsd./µl)",
      "Lymphocytes abs. (Tsd./µl)",
      "Monocytes abs. (Tsd./µl)",
      "Eosinophils abs. (Tsd./µl)",
      "Basophils abs. (Tsd./µl)",
      "Immature Granulocytes abs. (Tsd./µl)",
      "Platelets (Tsd./µl)"
    ) ~ "Circulating immune cells",

    TRUE ~ "Other"
  )
}

# ---- helper: apply manual marker order
apply_manual_marker_order <- function(df, variable_col = "Variable") {
  var_sym <- rlang::sym(variable_col)

  marker_levels_present <- df %>%
    distinct(!!var_sym) %>%
    mutate(
      Variable_clean = strip_dagger(as.character(!!var_sym)),
      order_id = match(Variable_clean, marker_order_clean)
    ) %>%
    arrange(is.na(order_id), order_id, !!var_sym) %>%
    pull(!!var_sym)

  df %>%
    mutate(marker = factor(!!var_sym, levels = rev(marker_levels_present)))
}

# ---- extract interaction term
extract_term_exact <- function(df, phase_label, term_exact, term_regex_fallback = NULL) {

  df2 <- df %>%
    mutate(
      Term = as.character(Term),
      Variable = as.character(Variable),
      Estimate = suppressWarnings(as.numeric(Estimate)),
      signif_star = as.character(`Signif.`)
    )

  out <- df2 %>% filter(Term == term_exact)

  if (nrow(out) == 0 && !is.null(term_regex_fallback)) {
    out <- df2 %>%
      filter(str_detect(Term, regex(term_regex_fallback, ignore_case = TRUE)))
  }

  if (nrow(out) == 0) {
    stop(
      paste0(
        "Could not find interaction term for phase='", phase_label, "'.\n",
        "Expected exact Term = '", term_exact, "'.\n",
        "Available Term values are:\n- ",
        paste(unique(df2$Term), collapse = "\n- ")
      )
    )
  }

  out %>% mutate(phase = phase_label)
}

fast_int <- extract_term_exact(
  fast_int_tab,
  phase_label = "Fasting",
  term_exact  = "Timepoint × fasting duration (Δ per day)",
  term_regex_fallback = "timepoint\\s*[×x]\\s*fasting\\s*duration.*(Δ|delta)\\s*per\\s*day"
)

refe_int <- extract_term_exact(
  refeed_int_tab,
  phase_label = "Refeeding",
  term_exact  = "Timepoint × refeeding duration (Δ per day)",
  term_regex_fallback = "timepoint\\s*[×x]\\s*refeeding\\s*duration.*(Δ|delta)\\s*per\\s*day"
)

# ---- combine
lme_int_all <- bind_rows(fast_int, refe_int) %>%
  mutate(
    phase = factor(phase, levels = c("Fasting", "Refeeding")),
    dose_label = if_else(phase == "Fasting", "Δ per fasting day", "Δ per refeeding day"),
    dose_label = factor(dose_label, levels = c("Δ per fasting day", "Δ per refeeding day")),
    domain = assign_domain_from_variable(Variable),
    domain = factor(domain, levels = c("Lipid", "Metabolic / Organ", "Inflammation & iron", "Circulating immune cells", "Other")),
    signif_star = if_else(is.na(signif_star) | signif_star == "ns", "", signif_star)
  )

# ---- manual marker order
lme_int_all <- apply_manual_marker_order(lme_int_all, variable_col = "Variable")

# ---- scaling
lims_int <- range(lme_int_all$Estimate, na.rm = TRUE)

signed_sqrt_trans <- scales::trans_new(
  name = "signed_sqrt",
  transform = function(x) sign(x) * sqrt(abs(x)),
  inverse   = function(x) sign(x) * (abs(x)^2)
)

breaks_int <- c(round(lims_int[1], 2), 0, round(lims_int[2], 2))

# ---- plot
p_heat_int <- ggplot(lme_int_all, aes(x = dose_label, y = marker, fill = Estimate)) +
  geom_tile(color = "grey92", linewidth = 0.25) +
  geom_text(aes(label = signif_star), size = 4.7) +
  facet_grid(
    domain ~ phase,
    scales = "free",
    space  = "free_y",
    switch = "y"
  ) +
  scale_y_discrete(expand = expansion(mult = 0.04)) +
  scale_x_discrete(drop = TRUE, expand = expansion(mult = 0.04)) +
  scale_fill_gradient2(
    low = "#0f766e",
    mid = "#ddd5c6",
    high = "#f59e0b",
    midpoint = 0,
    limits = lims_int,
    oob = scales::squish,
    trans = signed_sqrt_trans,
    breaks = breaks_int,
    name = "Interaction β\n(Δ per day)"
  ) +
  labs(x = NULL, y = NULL) +
  theme_bw() +
  theme(
    strip.placement = "outside",
    strip.background = element_blank(),
    strip.text.y.left = element_text(angle = 0, face = "bold", size = 11),
    strip.text.x = element_text(face = "bold", size = 12),
    panel.spacing.y = unit(0.5, "lines"),
    panel.spacing.x = unit(0.5, "lines"),
    panel.grid = element_blank(),
    plot.margin = margin(5, 5, 5, 5),
    axis.text.y = element_text(size = 10),
    axis.text.x = element_text(size = 11),
    legend.position = "right"
  )

p_heat_int

# ---- save
out_dir <- "C:/Users/megbi/Desktop/BA/Differential Blood Count Graphs/Master script 2.0/ResultsFigures"
dir.create(out_dir, recursive = TRUE, showWarnings = FALSE)

ggsave(file.path(out_dir, "heatmap_LME_interaction_deltaPerDay_two_columns_fixedOrder.png"),
       p_heat_int, width = 10.5, height = 10.5, dpi = 300)

ggsave(file.path(out_dir, "heatmap_LME_interaction_deltaPerDay_two_columns_fixedOrder.pdf"),
       p_heat_int, width = 10.5, height = 10.5)

summary(lme_int_all$Estimate)
```

## 7) ANCOVA Additive Heatmap

```{r ANOCVA-heatmap-additive, message=FALSE, warning=FALSE}
suppressPackageStartupMessages({
  library(tidyverse)
  library(readr)
  library(stringr)
  library(scales)
  library(forcats)
})

# ---- paths
base_dir <- "C:/Users/megbi/Desktop/BA/Differential Blood Count Graphs/Master script 2.0/ModelTables"
fast_main_path <- file.path(base_dir, "LM_fasting_ANCOVA_main_all_terms_T1onT0.csv")
ref_main_path  <- file.path(base_dir, "LM_refeed_ANCOVA_centered_all_terms_T2onT1.csv")

fast_main <- read_csv(fast_main_path, show_col_types = FALSE)
ref_main  <- read_csv(ref_main_path,  show_col_types = FALSE)

# ---- robust numeric parser
to_num <- function(x) suppressWarnings(as.numeric(gsub(",", ".", trimws(as.character(x)))))

# ---- additive-term extractor
extract_additive <- function(df, phase_label, term_regex, predictor_label) {
  df %>%
    mutate(
      Term = as.character(Term),
      Variable = as.character(Variable),
      Estimate = to_num(Estimate),
      signif_star = as.character(`Signif.`)
    ) %>%
    filter(str_detect(Term, regex(term_regex, ignore_case = TRUE))) %>%
    transmute(
      Variable,
      Term,
      Estimate,
      signif_star,
      phase = phase_label,
      predictor = predictor_label
    )
}

# ---- extract additive predictors
fast_bmi <- extract_additive(
  fast_main,
  phase_label = "Fasting",
  term_regex = "^(?!.*(T0|T1|:|×|x\\s)).*\\bBMI\\b",
  predictor_label = "BMI"
)

fast_dur <- extract_additive(
  fast_main,
  phase_label = "Fasting",
  term_regex = "^(?!.*(T0|T1|:|×)).*(fast|fasting).*dur|\\bfast\\w*\\b",
  predictor_label = "Fasting duration"
)

ref_bmi <- extract_additive(
  ref_main,
  phase_label = "Refeeding",
  term_regex = "^(?!.*(T1|T2|:|×|x\\s)).*\\bBMI\\b",
  predictor_label = "BMI"
)

ref_dur <- extract_additive(
  ref_main,
  phase_label = "Refeeding",
  term_regex = "^(?!.*(T1|T2|:|×)).*(refeed|refeeding|FR|reintro).*dur|\\bFR\\b",
  predictor_label = "Refeeding duration"
)

ref_diet <- extract_additive(
  ref_main,
  phase_label = "Refeeding",
  term_regex = "^\\s*Diet\\s*:",
  predictor_label = "Diet"
)

# ---- combine
ancova_add <- bind_rows(fast_bmi, fast_dur, ref_bmi, ref_dur, ref_diet) %>%
  distinct(Variable, phase, predictor, .keep_all = TRUE) %>%
  filter(!is.na(Estimate)) %>%
  mutate(
    domain = assign_domain_from_variable(Variable),
    domain = factor(
      domain,
      levels = c("Lipid", "Metabolic / Organ", "Inflammation & iron", "Circulating immune cells", "Other")
    ),
    phase = factor(phase, levels = c("Fasting", "Refeeding"))
  )

ancova_add <- bind_rows(fast_bmi, fast_dur, ref_bmi, ref_dur, ref_diet) %>%
  distinct(Variable, phase, predictor, .keep_all = TRUE) %>%
  filter(!is.na(Estimate)) %>%
  mutate(
    signif_star = case_when(
      is.na(signif_star) ~ "",
      signif_star == "ns" ~ "",
      TRUE ~ signif_star
    ),
    domain = assign_domain_from_variable(Variable),
    domain = factor(
      domain,
      levels = c("Lipid", "Metabolic / Organ", "Inflammation & iron", "Circulating immune cells", "Other")
    ),
    phase = factor(phase, levels = c("Fasting", "Refeeding"))
  )

# ---- manual marker order
ancova_add <- apply_manual_marker_order(ancova_add, variable_col = "Variable")

# ---- predictor order
predictor_order <- c("BMI", "Fasting duration", "Refeeding duration", "Diet")

ancova_add <- ancova_add %>%
  mutate(predictor = factor(predictor, levels = predictor_order)) %>%
  group_by(phase) %>%
  mutate(predictor = fct_drop(predictor)) %>%
  ungroup()

lims <- range(ancova_add$Estimate, na.rm = TRUE)
lims_pad <- c(lims[1], lims[2] + 1e-6)

signed_sqrt_trans <- scales::trans_new(
  name = "signed_sqrt",
  transform = function(x) sign(x) * sqrt(abs(x)),
  inverse   = function(x) sign(x) * (abs(x)^2)
)

breaks3 <- c(lims[1], 0, lims[2])
labels3 <- c(round(lims[1], 2), 0, round(lims[2], 2))

# ---- plot
p_add_heat <- ggplot(ancova_add, aes(x = predictor, y = marker, fill = Estimate)) +
  geom_tile(color = "grey92", linewidth = 0.25) +
  geom_text(aes(label = signif_star), size = 4.7) +
  facet_grid(
    domain ~ phase,
    scales = "free",
    space = "free_y",
    switch = "y"
  ) +
  scale_y_discrete(expand = expansion(mult = 0.04)) +
  scale_x_discrete(drop = TRUE, expand = expansion(mult = 0.04)) +
  scale_fill_gradient2(
  low = "#0f766e",
  mid = "#ddd5c6",
  high = "#f59e0b",
  midpoint = 0,
  limits = lims_pad,
  oob = scales::squish,
  trans = signed_sqrt_trans,
  breaks = breaks3,
  labels = labels3,
  name = "Additive β\n(model scale)"
) +
  labs(x = NULL, y = NULL) +
  theme_bw() +
  theme(
    strip.placement = "outside",
    strip.background = element_blank(),
    strip.text.y.left = element_text(angle = 0, face = "bold", size = 11),
    strip.text.x = element_text(face = "bold", size = 12),
    panel.spacing.y = unit(0.5, "lines"),
    panel.spacing.x = unit(0.5, "lines"),
    panel.grid = element_blank(),
    plot.margin = margin(5, 5, 5, 5),
    axis.text.y = element_text(size = 10),
    axis.text.x = element_text(size = 10),
    legend.position = "right"
  )

p_add_heat

# ---- save
out_dir <- "C:/Users/megbi/Desktop/BA/Differential Blood Count Graphs/Master script 2.0/ResultsFigures"
dir.create(out_dir, recursive = TRUE, showWarnings = FALSE)

ggsave(
  file.path(out_dir, "heatmap_ANCOVA_additive_predictors_fixedOrder.png"),
  p_add_heat, width = 12.5, height = 10.5, dpi = 300
)

ggsave(
  file.path(out_dir, "heatmap_ANCOVA_additive_predictors_fixedOrder.pdf"),
  p_add_heat, width = 12.5, height = 10.5
)

summary(ancova_add$Estimate)
```

## 8) ANCOVA Interaction Heatmap

```{r ANCOVA-heatmap-hetero-fixedOrder, message=FALSE, warning=FALSE}
suppressPackageStartupMessages({
  library(tidyverse)
  library(readr)
  library(stringr)
  library(scales)
  library(forcats)
  library(grid)
})

# ---- paths
base_dir <- "C:/Users/megbi/Desktop/BA/Differential Blood Count Graphs/Master script 2.0/ModelTables"

fast_bmi_path <- file.path(base_dir, "LM_fasting_ANCOVA_T0xBMI_all_terms_T1onT0.csv")
fast_dur_path <- file.path(base_dir, "LM_fasting_ANCOVA_T0xFast_all_terms_T1onT0.csv")

ref_bmi_path  <- file.path(base_dir, "LM_refeed_ANCOVA_T2onT1_BMIInteraction.csv")
ref_dur_path  <- file.path(base_dir, "LM_refeed_ANCOVA_T2onT1_T1xFR_centered_all_terms.csv")
ref_diet_path <- file.path(base_dir, "LM_refeed_ANCOVA_T2onT1_dietInteraction.csv")

# ---- read data
fast_bmi <- read_csv(fast_bmi_path, show_col_types = FALSE)
fast_dur <- read_csv(fast_dur_path, show_col_types = FALSE)

ref_bmi  <- read_csv(ref_bmi_path,  show_col_types = FALSE)
ref_dur  <- read_csv(ref_dur_path,  show_col_types = FALSE)
ref_diet <- read_csv(ref_diet_path, show_col_types = FALSE)

# ---- robust numeric parser
to_num <- function(x) suppressWarnings(as.numeric(gsub(",", ".", trimws(as.character(x)))))

# ---- extractor
extract_interaction <- function(df, phase, modulator, term_exact, term_regex_fallback = NULL) {
  
  df2 <- df %>%
    mutate(
      Term = as.character(Term),
      Variable = as.character(Variable),
      Estimate = to_num(Estimate),
      signif_star = as.character(`Signif.`)
    )
  
  out <- df2 %>% filter(Term == term_exact)
  
  if (nrow(out) == 0 && !is.null(term_regex_fallback)) {
    out <- df2 %>%
      filter(str_detect(Term, regex(term_regex_fallback, ignore_case = TRUE)))
  }
  
  if (nrow(out) == 0) {
    stop(
      paste0(
        "No rows found for phase='", phase, "', modulator='", modulator, "'.\n",
        "Expected Term: '", term_exact, "'.\n",
        "Available Terms:\n- ",
        paste(unique(df2$Term), collapse = "\n- ")
      )
    )
  }
  
  out %>%
    transmute(
      Variable,
      Term,
      Estimate,
      signif_star,
      phase = phase,
      modulator = modulator
    )
}

# ---- extract target interaction terms
fast_int_bmi <- extract_interaction(
  fast_bmi,
  phase = "Fasting",
  modulator = "BMI",
  term_exact = "T0 × BMI (baseline dependency differs by BMI)",
  term_regex_fallback = "T0\\s*[×x]\\s*BMI"
)

fast_int_dur <- extract_interaction(
  fast_dur,
  phase = "Fasting",
  modulator = "Fasting duration",
  term_exact = "T0 × fasting duration (baseline dependency differs by duration)",
  term_regex_fallback = "T0\\s*[×x]\\s*fasting\\s*duration|T0\\s*[×x]\\s*Fast"
)

ref_int_bmi <- extract_interaction(
  ref_bmi,
  phase = "Refeeding",
  modulator = "BMI",
  term_exact = "T1 × BMI (baseline dependency differs by BMI)",
  term_regex_fallback = "T1\\s*[×x]\\s*BMI"
)

ref_int_dur <- extract_interaction(
  ref_dur,
  phase = "Refeeding",
  modulator = "Refeeding duration",
  term_exact = "T1 × refeeding duration (baseline dependency mod.)",
  term_regex_fallback = "T1\\s*[×x]\\s*refeeding\\s*duration|T1\\s*[×x]\\s*FR"
)

ref_int_diet <- extract_interaction(
  ref_diet,
  phase = "Refeeding",
  modulator = "Diet",
  term_exact = "T1 × Diet (baseline dependency differs by diet)",
  term_regex_fallback = "T1\\s*[×x]\\s*Diet"
)

# ---- combine
ancova_hetero <- bind_rows(
  fast_int_bmi, fast_int_dur,
  ref_int_bmi, ref_int_dur, ref_int_diet
) %>%
  filter(!is.na(Estimate)) %>%
  mutate(
    domain = assign_domain_from_variable(Variable),
    domain = factor(
      domain,
      levels = c("Lipid", "Metabolic / Organ", "Inflammation & iron", "Circulating immune cells", "Other")
    ),
    phase = factor(phase, levels = c("Fasting", "Refeeding"))
  )

ancova_hetero <- bind_rows(
  fast_int_bmi, fast_int_dur,
  ref_int_bmi, ref_int_dur, ref_int_diet
) %>%
  filter(!is.na(Estimate)) %>%
  mutate(
    signif_star = case_when(
      is.na(signif_star) ~ "",
      str_trim(signif_star) == "ns" ~ "",
      TRUE ~ signif_star
    ),
    domain = assign_domain_from_variable(Variable),
    domain = factor(
      domain,
      levels = c("Lipid", "Metabolic / Organ", "Inflammation & iron", "Circulating immune cells", "Other")
    ),
    phase = factor(phase, levels = c("Fasting", "Refeeding"))
  )


# ---- manual marker order
ancova_hetero <- apply_manual_marker_order(ancova_hetero, variable_col = "Variable")

# ---- phase-specific x-axis labels
ancova_hetero <- ancova_hetero %>%
  mutate(
    modulator = factor(modulator, levels = c("BMI", "Fasting duration", "Refeeding duration", "Diet")),
    modulator_label = case_when(
      phase == "Fasting"   & modulator == "BMI"                ~ "T0 × BMI",
      phase == "Fasting"   & modulator == "Fasting duration"   ~ "T0 ×\nFasting duration",
      phase == "Refeeding" & modulator == "BMI"                ~ "T1 × BMI",
      phase == "Refeeding" & modulator == "Refeeding duration" ~ "T1 ×\nRefeeding duration",
      phase == "Refeeding" & modulator == "Diet"               ~ "T1 × Diet",
      TRUE ~ as.character(modulator)
    )
  ) %>%
  group_by(phase) %>%
  mutate(modulator_label = factor(modulator_label, levels = unique(modulator_label))) %>%
  ungroup()

# ---- scaling: actual observed min/max
lims <- range(ancova_hetero$Estimate, na.rm = TRUE)

signed_sqrt_trans <- scales::trans_new(
  name = "signed_sqrt",
  transform = function(x) sign(x) * sqrt(abs(x)),
  inverse   = function(x) sign(x) * (abs(x)^2)
)

breaks3 <- c(round(lims[1], 2), 0, round(lims[2], 2))

# ---- plot
p_hetero_heat <- ggplot(ancova_hetero, aes(x = modulator_label, y = marker, fill = Estimate)) +
  geom_tile(color = "grey92", linewidth = 0.25) +
  geom_text(aes(label = signif_star), size = 4.7) +
  facet_grid(
    domain ~ phase,
    scales = "free",
    space = "free_y",
    switch = "y"
  ) +
  scale_y_discrete(expand = expansion(mult = 0.04)) +
  scale_x_discrete(drop = TRUE, expand = expansion(mult = 0.08)) +
  scale_fill_gradient2(
    low = "#0f766e",
    mid = "#ddd5c6",
    high = "#f59e0b",
    midpoint = 0,
    limits = lims,
    oob = scales::squish,
    trans = signed_sqrt_trans,
    breaks = breaks3,
    name = "Interaction β\n(model scale)"
  ) +
  labs(x = NULL, y = NULL) +
  theme_bw() +
  theme(
    strip.placement = "outside",
    strip.background = element_blank(),
    strip.text.y.left = element_text(angle = 0, face = "bold", size = 11),
    strip.text.x = element_text(face = "bold", size = 12),
    panel.spacing.x = unit(0.8, "lines"),
    panel.spacing.y = unit(0.5, "lines"),
    panel.grid = element_blank(),
    plot.margin = margin(5, 12, 5, 5),
    axis.text.y = element_text(size = 10),
    axis.text.x = element_text(size = 9, lineheight = 0.9, margin = margin(t = 4)),
    legend.position = "right"
  )

p_hetero_heat

# ---- save
out_dir <- "C:/Users/megbi/Desktop/BA/Differential Blood Count Graphs/Master script 2.0/ResultsFigures"
dir.create(out_dir, recursive = TRUE, showWarnings = FALSE)

ggsave(
  file.path(out_dir, "heatmap_ANCOVA_heterogeneity_interactions_fixedOrder.png"),
  p_hetero_heat, width = 12.5, height = 10.5, dpi = 300
)

ggsave(
  file.path(out_dir, "heatmap_ANCOVA_heterogeneity_interactions_fixedOrder.pdf"),
  p_hetero_heat, width = 12.5, height = 10.5
)

summary(ancova_hetero$Estimate)
```

## 12) Immune-inflammatory indices graph

```{r immune-indices-all-in-one, message=FALSE, warning=FALSE}
# ============================================================
# Immune indices: calculate ratios, fit LME phase contrasts,
# output results table, and plot trajectories with significance
# ============================================================

suppressPackageStartupMessages({
  library(dplyr)
  library(tidyr)
  library(ggplot2)
  library(lme4)
  library(lmerTest)
  library(emmeans)
  library(knitr)
  library(kableExtra)
  library(readr)
})

# -----------------------------
# Settings
# -----------------------------
min_n <- 20
index_vars <- c("LMR (ratio)", "NLR (ratio)", "PLR (ratio)", "SII (a.u.)")


# ============================================================
# 1. Calculate immune indices from absolute counts
# ============================================================

# Keep only the markers needed for index calculation
idx_source <- FF_laboratory_master %>%
  filter(
    variable %in% c(
      "Lymphocytes abs. (Tsd./µl)",
      "Monocytes abs. (Tsd./µl)",
      "Neutrophils abs. (Tsd./µl)",
      "Platelets (Tsd./µl)"
    ),
    timepoint %in% c("T0", "T1", "T2")
  ) %>%
  group_by(study_id, variable, timepoint) %>%
  slice_min(datetime, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  transmute(
    study_id,
    timepoint = factor(timepoint, levels = c("T0", "T1", "T2")),
    datetime,
    variable,
    value = suppressWarnings(as.numeric(value)),
    age = suppressWarnings(as.numeric(age)),
    sex = factor(sex, levels = c("F", "M")),
    BMI_baseline = suppressWarnings(as.numeric(BMI_baseline))
  )

# Pivot wide for ratio calculation
idx_wide <- idx_source %>%
  select(study_id, timepoint, datetime, variable, value, age, sex, BMI_baseline) %>%
  pivot_wider(
    names_from = variable,
    values_from = value
  )

# Calculate indices using absolute counts
immune_indices_long <- idx_wide %>%
  mutate(
    `LMR (ratio)` = `Lymphocytes abs. (Tsd./µl)` / `Monocytes abs. (Tsd./µl)`,
    `NLR (ratio)` = `Neutrophils abs. (Tsd./µl)` / `Lymphocytes abs. (Tsd./µl)`,
    `PLR (ratio)` = `Platelets (Tsd./µl)` / `Lymphocytes abs. (Tsd./µl)`,
    `SII (a.u.)`  = (`Platelets (Tsd./µl)` * `Neutrophils abs. (Tsd./µl)`) / `Lymphocytes abs. (Tsd./µl)`
  ) %>%
  mutate(
    across(all_of(index_vars), ~ ifelse(is.finite(.x) & .x > 0, .x, NA_real_))
  ) %>%
  select(
    study_id, timepoint, datetime, age, sex, BMI_baseline,
    all_of(index_vars)
  ) %>%
  pivot_longer(
    cols = all_of(index_vars),
    names_to = "index",
    values_to = "value"
  ) %>%
  filter(!is.na(value), value > 0)

# Optional quick check
immune_index_counts <- immune_indices_long %>%
  count(index, timepoint)

print(immune_index_counts)

# ============================================================
# 2. LME phase-specific inference (T1-T0, T2-T1)
# ============================================================

fit_one_index <- function(df, idx_name) {

  d <- df %>%
    filter(index == idx_name) %>%
    filter(!is.na(value), value > 0) %>%
    filter(!is.na(age), !is.na(sex), !is.na(BMI_baseline))

  n_subj <- n_distinct(d$study_id)

  if (n_subj < min_n) {
    return(tibble(
      index = idx_name,
      Phase = c("Fasting", "Refeeding"),
      ratio = NA_real_,
      pct_change = NA_real_,
      ratio_ci_low = NA_real_,
      ratio_ci_high = NA_real_,
      p_raw = NA_real_,
      n_subjects = n_subj
    ))
  }

  d <- d %>%
    mutate(
      age_c = age - mean(age, na.rm = TRUE),
      BMI_c = BMI_baseline - mean(BMI_baseline, na.rm = TRUE),
      y = log(value)
    )

  fit <- lmer(y ~ timepoint + age_c + sex + BMI_c + (1 | study_id), data = d)

  em <- emmeans(fit, ~ timepoint)

  ctr <- contrast(
    em,
    method = list(
      "Fasting"   = c(-1,  1,  0),  # T1 - T0
      "Refeeding" = c( 0, -1,  1)   # T2 - T1
    ),
    adjust = "none"
  )

  ctr_df <- as.data.frame(summary(ctr, infer = c(TRUE, TRUE)))

  # robust CI naming across emmeans versions
  if (!("lower.CL" %in% names(ctr_df)) && ("asymp.LCL" %in% names(ctr_df))) {
    ctr_df <- ctr_df %>% rename(lower.CL = asymp.LCL, upper.CL = asymp.UCL)
  }
  if (!("lower.CL" %in% names(ctr_df))) ctr_df$lower.CL <- NA_real_
  if (!("upper.CL" %in% names(ctr_df))) ctr_df$upper.CL <- NA_real_

  ctr_df %>%
    transmute(
      index = idx_name,
      Phase = contrast,
      ratio = exp(estimate),
      pct_change = (exp(estimate) - 1) * 100,
      ratio_ci_low = exp(lower.CL),
      ratio_ci_high = exp(upper.CL),
      p_raw = p.value,
      n_subjects = n_subj
    )
}

phase_res <- bind_rows(lapply(index_vars, function(idx) fit_one_index(immune_indices_long, idx))) %>%
  mutate(p_fdr = p.adjust(p_raw, method = "BH"))

# Pretty results table
results_pretty <- phase_res %>%
  mutate(
    Index = case_match(index,
  "LMR (ratio)" ~ "LMR", "NLR (ratio)" ~ "NLR",
  "PLR (ratio)" ~ "PLR", "SII (a.u.)"  ~ "SII (a.u.)",
  .default = index),
    `% change` = round(pct_change, 1),
    `95% CI (post/pre)` = ifelse(
      is.na(ratio_ci_low) | is.na(ratio_ci_high),
      NA_character_,
      paste0("[", round(ratio_ci_low, 3), ", ", round(ratio_ci_high, 3), "]")
    ),
    `P (raw)` = fmt_p(p_raw),
    `P (FDR)` = fmt_p(p_fdr),
    `Signif.` = p_stars(p_fdr)
  ) %>%
  select(
    Index,
    Phase,
    `% change`,
    `95% CI (post/pre)`,
    `P (raw)`,
    `P (FDR)`,
    `Signif.`,
    `n (subjects)` = n_subjects
  ) %>%
  arrange(Index, factor(Phase, levels = c("Fasting", "Refeeding")))

results_pretty %>%
  kbl(
    caption = "Phase-specific changes of immune indices. LME on log(index) with covariates.",
    align = "l"
  ) %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE) %>%
  scroll_box(width = "100%", height = "520px")

# Save results table
dir.create("ModelTables", showWarnings = FALSE)
write_csv(results_pretty, "ModelTables/LME_immune_indices_phase_table_T1T0_T2T1.csv")

# ============================================================
# 3. Plot means ± 95% CI with significance brackets
# ============================================================

plot_df <- immune_indices_long %>%
  group_by(index, timepoint) %>%
  summarise(
    n = sum(!is.na(value)),
    mean = mean(value, na.rm = TRUE),
    sd = sd(value, na.rm = TRUE),
    se = sd / sqrt(n),
    ci_low = mean - 1.96 * se,
    ci_high = mean + 1.96 * se,
    .groups = "drop"
  ) %>%
  mutate(
    Index = case_match(index,
  "LMR (ratio)" ~ "LMR", "NLR (ratio)" ~ "NLR",
  "PLR (ratio)" ~ "PLR", "SII (a.u.)"  ~ "SII (a.u.)",
  .default = index),
    Timepoint = factor(timepoint, levels = c("T0", "T1", "T2"))
  )

sig_df <- phase_res %>%
  mutate(
    Index = case_match(index,
  "LMR (ratio)" ~ "LMR", "NLR (ratio)" ~ "NLR",
  "PLR (ratio)" ~ "PLR", "SII (a.u.)"  ~ "SII (a.u.)",
  .default = index),
    stars = case_when(
      is.na(p_fdr) ~ "",
      p_fdr < 0.001 ~ "☆☆☆",
      p_fdr < 0.01  ~ "☆☆",
      p_fdr < 0.05  ~ "☆",
      TRUE ~ "ns"
    ),
    x_start = case_when(
      Phase == "Fasting" ~ 1,
      Phase == "Refeeding" ~ 2
    ),
    x_end = case_when(
      Phase == "Fasting" ~ 2,
      Phase == "Refeeding" ~ 3
    )
  )

panel_heights <- plot_df %>%
  group_by(Index) %>%
  summarise(
    ymax = max(ci_high, na.rm = TRUE),
    ymin = min(ci_low, na.rm = TRUE),
    yrange = ymax - ymin,
    .groups = "drop"
  )

sig_df <- sig_df %>%
  left_join(panel_heights, by = "Index") %>%
  mutate(
    bracket_y = case_when(
      Phase == "Fasting"   ~ ymax + 0.10 * yrange,
      Phase == "Refeeding" ~ ymax + 0.22 * yrange
    ),
    label_y = bracket_y + 0.04 * yrange,
    tip_y = bracket_y - 0.03 * yrange
  )

p <- ggplot(plot_df, aes(x = Timepoint, y = mean, group = 1)) +
  geom_line(color = "black") +
  geom_point(size = 2.8, color = "black") +
  geom_errorbar(aes(ymin = ci_low, ymax = ci_high), width = 0.08) +
  facet_wrap(~ Index, scales = "free_y") +
  geom_segment(
    data = sig_df,
    aes(x = x_start, xend = x_end, y = bracket_y, yend = bracket_y),
    inherit.aes = FALSE
  ) +
  geom_segment(
    data = sig_df,
    aes(x = x_start, xend = x_start, y = bracket_y, yend = tip_y),
    inherit.aes = FALSE
  ) +
  geom_segment(
    data = sig_df,
    aes(x = x_end, xend = x_end, y = bracket_y, yend = tip_y),
    inherit.aes = FALSE
  ) +
  geom_text(
    data = sig_df,
    aes(x = (x_start + x_end) / 2, y = label_y, label = stars),
    inherit.aes = FALSE,
    size = 3.2
  ) +
  labs(
    x = "Timepoint",
    y = "Mean ± 95% CI"
  ) +
  theme_bw(base_size = 12) +
  theme(
    strip.background = element_rect(fill = "grey85"),
    strip.text = element_text(face = "bold"),
    axis.title.x = element_text(margin = margin(t = 8)),
    axis.title.y = element_text(margin = margin(r = 8))
  )

p

# Save figure
dir.create("Figures", showWarnings = FALSE)
ggsave("Figures/immune_indices_trajectory.pdf", plot = p, width = 8, height = 6, units = "in")
ggsave("Figures/immune_indices_trajectory.png", plot = p, width = 8, height = 6, units = "in", dpi = 600)

# descriptive II graph numbers
immune_index_descriptives <- plot_df %>%
  select(Index, Timepoint, n, mean, ci_low, ci_high) %>%
  mutate(
    mean = round(mean, 2),
    ci_low = round(ci_low, 2),
    ci_high = round(ci_high, 2),
    `Mean ± 95% CI` = paste0(mean, " [", ci_low, ", ", ci_high, "]")
  ) %>%
  select(Index, Timepoint, n, `Mean ± 95% CI`, mean, ci_low, ci_high) %>%
  arrange(Index, Timepoint)

print(immune_index_descriptives)
immune_index_descriptives %>%
  knitr::kable(
    caption = "Descriptive summary of immune redistribution indices by timepoint.",
    align = "lcccccc"
  ) %>%
  kableExtra::kable_styling(
    bootstrap_options = c("striped", "hover"),
    full_width = FALSE
  )

write_csv(
  immune_index_descriptives,
  "ModelTables/immune_indices_descriptive_mean_ci_n.csv"
)
```

## 13) Immune indices diet heterogeneity

```{r immune-refeed, message=FALSE, warning=FALSE}
## ============================================================
## X) Immune indices and refeeding (T1 -> T2) by diet
## - Computes indices from absolute counts (earliest per subject/timepoint)
## - Builds refeeding dataset (T1, T2)
## - Plots T1 vs Δ Index (T2 - T1) by diet in a 2x2 panel
## - Panel order matches the average trajectories graph:
##   LMR, NLR, PLR, SII (a.u.)
## ============================================================

suppressPackageStartupMessages({
  library(dplyr)
  library(tidyr)
  library(ggplot2)
})

## -------------------------
## 0) Inputs
## -------------------------
index_vars_needed <- c(
  "Neutrophils abs. (Tsd./µl)",
  "Lymphocytes abs. (Tsd./µl)",
  "Monocytes abs. (Tsd./µl)",
  "Platelets (Tsd./µl)"
)

panel_order <- c("LMR", "NLR", "PLR", "SII (a.u.)")

index_ranges <- tibble::tribble(
  ~Index,        ~xmin,  ~xmax,
  "LMR",           3.0,   5.5,
  "NLR",           1.0,   3.0,
  "PLR",         80.0, 160.0,
  "SII (a.u.)",  300.0, 700.0
)

## -------------------------
## 1) Build wide table of components (earliest per subject/timepoint)
## -------------------------
index_wide <- FF_laboratory_master %>%
  filter(
    timepoint %in% c("T0", "T1", "T2"),
    variable %in% index_vars_needed
  ) %>%
  group_by(study_id, timepoint, variable) %>%
  slice_min(datetime, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  select(
    study_id, sex, timepoint, datetime,
    fasting_duration_measured, FR_duration,
    age, BMI_baseline, diet_keto,
    variable, value
  ) %>%
  pivot_wider(names_from = variable, values_from = value)

## -------------------------
## 2) Compute indices safely
## -------------------------
indexes_long <- index_wide %>%
    mutate(
        NLR = if_else(
            !is.na(`Neutrophils abs. (Tsd./µl)`) &
                !is.na(`Lymphocytes abs. (Tsd./µl)`) &
                `Lymphocytes abs. (Tsd./µl)` > 0,
            `Neutrophils abs. (Tsd./µl)` / `Lymphocytes abs. (Tsd./µl)`,
            NA_real_
        ),
        PLR = if_else(
            !is.na(`Platelets (Tsd./µl)`) &
                !is.na(`Lymphocytes abs. (Tsd./µl)`) &
                `Lymphocytes abs. (Tsd./µl)` > 0,
            `Platelets (Tsd./µl)` / `Lymphocytes abs. (Tsd./µl)`,
            NA_real_
        ),
        LMR = if_else(
            !is.na(`Lymphocytes abs. (Tsd./µl)`) &
                !is.na(`Monocytes abs. (Tsd./µl)`) &
                `Monocytes abs. (Tsd./µl)` > 0,
            `Lymphocytes abs. (Tsd./µl)` / `Monocytes abs. (Tsd./µl)`,
            NA_real_
        ),
        SII = if_else(
            !is.na(`Platelets (Tsd./µl)`) &
                !is.na(`Neutrophils abs. (Tsd./µl)`) &
                !is.na(`Lymphocytes abs. (Tsd./µl)`) &
                `Lymphocytes abs. (Tsd./µl)` > 0,
            `Platelets (Tsd./µl)` * `Neutrophils abs. (Tsd./µl)` / `Lymphocytes abs. (Tsd./µl)`,
            NA_real_
        )
    ) %>%
    select(
        study_id, sex, timepoint, datetime,
        fasting_duration_measured, FR_duration,
        age, BMI_baseline, diet_keto,
        NLR, PLR, LMR, SII
    ) %>%
    pivot_longer(
        cols = c(LMR, NLR, PLR, SII),
        names_to = "Index",
        values_to = "value"
    ) %>%
    mutate(
        Index = dplyr::recode(
            Index,
            LMR = "LMR",
            NLR = "NLR",
            PLR = "PLR",
            SII = "SII (a.u.)"
        ),
        Index = factor(Index, levels = panel_order)
    )

## Optional: bind into master if you still want that in your workflow
FF_laboratory_master <- bind_rows(
  FF_laboratory_master,
  indexes_long %>%
    transmute(
      study_id, sex, timepoint, datetime,
      fasting_duration_measured, FR_duration,
      age, BMI_baseline, diet_keto,
      variable = as.character(Index),
      value
    )
)

## -------------------------
## 3) Refeeding dataset (T1, T2) with diet
## -------------------------
delta_indexes_refeed <- indexes_long %>%
  filter(!is.na(diet_keto)) %>%
  filter(timepoint %in% c("T1", "T2")) %>%
  group_by(study_id, Index, timepoint) %>%
  slice_min(datetime, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  pivot_wider(
    id_cols = c(study_id, Index, sex, age, BMI_baseline, diet_keto, FR_duration),
    names_from = timepoint,
    values_from = value
  ) %>%
  mutate(
    delta_T2_T1 = T2 - T1,
    diet_keto = factor(diet_keto)
  ) %>%
  filter(!is.na(T1), !is.na(T2))

## n subjects per index
n_refeed_by_var <- delta_indexes_refeed %>%
  group_by(Index) %>%
  summarise(n_subjects = n_distinct(study_id), .groups = "drop")

print(n_refeed_by_var)

## -------------------------
## 4) Healthy band panel plot: T1 vs Δ Index (T2 - T1) by diet
## -------------------------
p_refeed_indices_panel <- ggplot(
  delta_indexes_refeed,
  aes(x = T1, y = delta_T2_T1, color = diet_keto)
) +
  geom_rect(
    data = index_ranges,
    aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf),
    inherit.aes = FALSE,
    fill = "darkseagreen3",
    alpha = 0.45
  ) +
  geom_point(alpha = 0.6, size = 1.8) +
  geom_smooth(method = "lm", se = TRUE, linewidth = 0.8) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  facet_wrap(~ Index, scales = "free", ncol = 2) +
  labs(
    x = "Index at end of fasting (T1)",
    y = expression(Delta~"Index (T2 - T1)"),
    color = "Diet"
  ) +
  theme_bw(base_size = 12) +
  theme(
    strip.background = element_rect(fill = "grey85"),
    strip.text = element_text(face = "bold"),
    axis.title.x = element_text(margin = margin(t = 8)),
    axis.title.y = element_text(margin = margin(r = 8))
  )

p_refeed_indices_panel

## -------------------------
## 5) Save figure
## -------------------------
dir.create("Figures", showWarnings = FALSE)

ggsave(
  "Figures/refeeding_immune_indices_by_diet_panel.pdf",
  plot = p_refeed_indices_panel,
  width = 8,
  height = 6,
  units = "in"
)

ggsave(
  "Figures/refeeding_immune_indices_by_diet_panel.png",
  plot = p_refeed_indices_panel,
  width = 8,
  height = 6,
  units = "in",
  dpi = 600
)

## ============================================================
## X) Immune indices and refeeding (T1 -> T2): ANCOVA models
## - Additive model:    log(T2) ~ T1_c + diet + FR + age + sex + BMI
## - Interaction model: log(T2) ~ T1_c * diet + FR + age + sex + BMI
## - Uses explicit complete-case filtering
## ============================================================

## -------------------------
## 1) Analysis dataset
## -------------------------
ii_refeed_ancova_df <- delta_indexes_refeed %>%
  filter(!is.na(T1), !is.na(T2)) %>%
  mutate(
    diet_keto = factor(diet_keto, levels = c("Non-Keto", "Keto"))
  ) %>%
  filter(T1 > 0, T2 > 0) %>%
  mutate(
    T1_log = log(T1),
    T2_log = log(T2),
    T1_c  = as.numeric(scale(T1_log, center = TRUE, scale = FALSE)),
    age_c = as.numeric(scale(age, center = TRUE, scale = FALSE)),
    BMI_c = as.numeric(scale(BMI_baseline, center = TRUE, scale = FALSE)),
    FR_c  = as.numeric(scale(FR_duration, center = TRUE, scale = FALSE))
  ) %>%
  drop_na(T1_c, T2_log, diet_keto, age_c, sex, BMI_c, FR_c)

## sanity check
dropped_refeed <- delta_indexes_refeed %>%
  summarise(
    n_total = n(),
    n_nonpos = sum((!is.na(T1) & T1 <= 0) | (!is.na(T2) & T2 <= 0), na.rm = TRUE)
  )

print(dropped_refeed)

n_refeed_model <- ii_refeed_ancova_df %>%
  group_by(Index) %>%
  summarise(n_subjects = n_distinct(study_id), .groups = "drop")

print(n_refeed_model)

## -------------------------
## 2) Additive ANCOVA
##    log(T2) ~ T1_c + diet + FR + age + sex + BMI
## -------------------------
ii_refeed_additive_results <- ii_refeed_ancova_df %>%
  group_by(Index) %>%
  group_modify(~{
    df <- .x

    fit <- lm(
      T2_log ~ T1_c + diet_keto + age_c + sex + BMI_c + FR_c,
      data = df
    )

    broom::tidy(fit, conf.int = TRUE) %>%
      filter(term %in% c("T1_c", "diet_ketoKeto", "FR_c", "age_c", "sexM", "BMI_c")) %>%
      mutate(
        n_rows = nrow(df),
        n_subjects = n_distinct(df$study_id)
      )
  }) %>%
  ungroup()

ii_refeed_additive_table <- ii_refeed_additive_results %>%
  group_by(term) %>%
  mutate(
    p_raw = p.value,
    p_fdr = p.adjust(p_raw, method = "BH")
  ) %>%
  ungroup() %>%
  mutate(
    Term = dplyr::recode(
      term,
      T1_c = "End of fasting (T1), centered",
      diet_ketoKeto = "Diet: Keto (vs Non-Keto)",
      FR_c = "Refeeding duration, centered",
      age_c = "Age, centered",
      sexM = "Sex: Male (vs Female)",
      BMI_c = "Baseline BMI, centered"
    ),
    Estimate = round(estimate, 3),
    SE = round(std.error, 3),
    `95% CI` = paste0("[", round(conf.low, 3), ", ", round(conf.high, 3), "]"),
    `P (raw)` = fmt_p(p_raw),
    `P (FDR)` = fmt_p(p_fdr),
    `Signif.` = p_stars(p_fdr)
  ) %>%
  select(
    Index,
    Term,
    Estimate,
    SE,
    `95% CI`,
    `P (raw)`,
    `P (FDR)`,
    `Signif.`,
    n_subjects
  ) %>%
  arrange(
    factor(Index, levels = c("LMR", "NLR", "PLR", "SII (a.u.)")),
    factor(
      Term,
      levels = c(
        "End of fasting (T1), centered",
        "Diet: Keto (vs Non-Keto)",
        "Refeeding duration, centered",
        "Baseline BMI, centered",
        "Age, centered",
        "Sex: Male (vs Female)"
      )
    )
  )

print(ii_refeed_additive_table, n = 24)

ii_refeed_additive_table %>%
  kbl(
    caption = "Refeeding ANCOVA for immune indices: additive model.",
    align = "l"
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover"),
    full_width = FALSE
  )

## -------------------------
## 3) Diet-interaction ANCOVA
##    log(T2) ~ T1_c * diet + FR + age + sex + BMI
## -------------------------
ii_refeed_interaction_results <- ii_refeed_ancova_df %>%
  group_by(Index) %>%
  group_modify(~{
    df <- .x

    fit <- lm(
      T2_log ~ T1_c * diet_keto + age_c + sex + BMI_c + FR_c,
      data = df
    )

    broom::tidy(fit, conf.int = TRUE) %>%
      filter(term %in% c("T1_c", "diet_ketoKeto", "T1_c:diet_ketoKeto", "FR_c", "age_c", "sexM", "BMI_c")) %>%
      mutate(
        n_rows = nrow(df),
        n_subjects = n_distinct(df$study_id)
      )
  }) %>%
  ungroup()

ii_refeed_interaction_table <- ii_refeed_interaction_results %>%
  group_by(term) %>%
  mutate(
    p_raw = p.value,
    p_fdr = p.adjust(p_raw, method = "BH")
  ) %>%
  ungroup() %>%
  mutate(
    Term = dplyr::recode(
      term,
      T1_c = "End of fasting (T1), centered",
      diet_ketoKeto = "Diet: Keto (vs Non-Keto)",
      `T1_c:diet_ketoKeto` = "T1 × Diet (baseline dependency differs by diet)",
      FR_c = "Refeeding duration, centered",
      age_c = "Age, centered",
      sexM = "Sex: Male (vs Female)",
      BMI_c = "Baseline BMI, centered"
    ),
    Estimate = round(estimate, 3),
    SE = round(std.error, 3),
    `95% CI` = paste0("[", round(conf.low, 3), ", ", round(conf.high, 3), "]"),
    `P (raw)` = fmt_p(p_raw),
    `P (FDR)` = fmt_p(p_fdr),
    `Signif.` = p_stars(p_fdr)
  ) %>%
  select(
    Index,
    Term,
    Estimate,
    SE,
    `95% CI`,
    `P (raw)`,
    `P (FDR)`,
    `Signif.`,
    n_subjects
  ) %>%
  arrange(
    factor(Index, levels = c("LMR", "NLR", "PLR", "SII (a.u.)")),
    factor(
      Term,
      levels = c(
        "T1 × Diet (baseline dependency differs by diet)",
        "End of fasting (T1), centered",
        "Diet: Keto (vs Non-Keto)",
        "Refeeding duration, centered",
        "Baseline BMI, centered",
        "Age, centered",
        "Sex: Male (vs Female)"
      )
    )
  )

print(ii_refeed_interaction_table, n = 28)

ii_refeed_interaction_table %>%
  kbl(
    caption = "Refeeding ANCOVA for immune indices: diet-interaction model.",
    align = "l"
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover"),
    full_width = FALSE
  )

## -------------------------
## 4) Save tables
## -------------------------
dir.create("ModelTables", showWarnings = FALSE)

write_csv(
  ii_refeed_additive_table,
  "ModelTables/ANCOVA_refeed_immune_indices_additive.csv"
)

write_csv(
  ii_refeed_interaction_table,
  "ModelTables/ANCOVA_refeed_immune_indices_diet_interaction.csv"
)
```

## 14) Baseline Inflammation and metabolic health

```{r Baseline Inflammation, message=FALSE, warning=FALSE}
# =========================================================
# Question:
# Does baseline CRP predict fasting-endpoint (T1) metabolic markers,
# adjusted for T0 and covariates?
# =========================================================
center_only <- function(x) {
  x - mean(x, na.rm = TRUE)
}

# -------------------------
# 1) Build baseline CRP predictor (T0)
# -------------------------
crp_T0 <- FF_laboratory_master %>%
  dplyr::filter(timepoint == "T0", variable == "CRP (mg/l)") %>%
  dplyr::group_by(study_id) %>%
  dplyr::slice_min(datetime, n = 1, with_ties = FALSE) %>%
  dplyr::ungroup() %>%
  dplyr::transmute(
    study_id,
    CRP_T0 = value,
    CRP_T0_log = safe_log(value)
  ) %>%
  dplyr::filter(!is.na(CRP_T0_log)) %>%
  dplyr::mutate(
    CRP_T0_z = as.numeric(scale(CRP_T0_log))
  )

# -------------------------
# 2) Build T0/T1 metabolic dataset
# -------------------------
met_vars_needed <- c(
  "Serum Glucose (mg/dl)",
  "Insulin (mIU/l)",
  "Triglycerides (mg/dl)",
  "HDL-C (mg/dl)",
  "LDL-C (mg/dl)",
  "Non-HDL Cholesterol (mg/dl)"
)

met_T0T1_wide <- FF_laboratory_master %>%
  dplyr::filter(timepoint %in% c("T0", "T1"), variable %in% met_vars_needed) %>%
  dplyr::group_by(study_id, timepoint, variable) %>%
  dplyr::slice_min(datetime, n = 1, with_ties = FALSE) %>%
  dplyr::ungroup() %>%
  dplyr::select(
    study_id, timepoint, variable, value,
    age, sex, BMI_baseline, fasting_duration_measured
  ) %>%
  tidyr::pivot_wider(names_from = variable, values_from = value) %>%
  tidyr::pivot_wider(
    names_from = timepoint,
    values_from = c(
      `Serum Glucose (mg/dl)`,
      `Insulin (mIU/l)`,
      `Triglycerides (mg/dl)`,
      `HDL-C (mg/dl)`,
      `LDL-C (mg/dl)`,
      `Non-HDL Cholesterol (mg/dl)`
    )
  ) %>%
  dplyr::mutate(
    # Rename raw markers
    Glucose_T0 = `Serum Glucose (mg/dl)_T0`,
    Glucose_T1 = `Serum Glucose (mg/dl)_T1`,
    Insulin_T0 = `Insulin (mIU/l)_T0`,
    Insulin_T1 = `Insulin (mIU/l)_T1`,
    TG_T0      = `Triglycerides (mg/dl)_T0`,
    TG_T1      = `Triglycerides (mg/dl)_T1`,
    HDL_T0     = `HDL-C (mg/dl)_T0`,
    HDL_T1     = `HDL-C (mg/dl)_T1`,
    LDL_T0     = `LDL-C (mg/dl)_T0`,
    LDL_T1     = `LDL-C (mg/dl)_T1`,
    nonHDL_T0  = `Non-HDL Cholesterol (mg/dl)_T0`,
    nonHDL_T1  = `Non-HDL Cholesterol (mg/dl)_T1`,

    # Derived indices
    TyG_T0 = ifelse(!is.na(TG_T0) & !is.na(Glucose_T0) & TG_T0 > 0 & Glucose_T0 > 0,
                    log((TG_T0 * Glucose_T0) / 2), NA_real_),
    TyG_T1 = ifelse(!is.na(TG_T1) & !is.na(Glucose_T1) & TG_T1 > 0 & Glucose_T1 > 0,
                    log((TG_T1 * Glucose_T1) / 2), NA_real_),

    AIP_T0 = ifelse(!is.na(TG_T0) & !is.na(HDL_T0) & TG_T0 > 0 & HDL_T0 > 0,
                    log10(TG_T0 / HDL_T0), NA_real_),
    AIP_T1 = ifelse(!is.na(TG_T1) & !is.na(HDL_T1) & TG_T1 > 0 & HDL_T1 > 0,
                    log10(TG_T1 / HDL_T1), NA_real_),

    HOMA_T0 = ifelse(!is.na(Glucose_T0) & !is.na(Insulin_T0) & Glucose_T0 > 0 & Insulin_T0 > 0,
                     ((Glucose_T0 / 18) * Insulin_T0) / 22.5, NA_real_),
    HOMA_T1 = ifelse(!is.na(Glucose_T1) & !is.na(Insulin_T1) & Glucose_T1 > 0 & Insulin_T1 > 0,
                     ((Glucose_T1 / 18) * Insulin_T1) / 22.5, NA_real_)
  ) %>%
  dplyr::left_join(
    crp_T0 %>% dplyr::select(study_id, CRP_T0_z),
    by = "study_id"
  )

# -------------------------
# 3) Long ANCOVA dataset (ALL outcomes)
# -------------------------
all_outcomes <- c("Glucose", "Insulin", "HOMA", "TyG", "AIP", "HDL", "LDL", "nonHDL")

ancova_long_crp <- met_T0T1_wide %>%
  dplyr::transmute(
    study_id, age, sex, BMI_baseline, fasting_duration_measured, CRP_T0_z,
    Glucose_T0, Glucose_T1,
    Insulin_T0, Insulin_T1,
    HOMA_T0, HOMA_T1,
    TyG_T0, TyG_T1,
    AIP_T0, AIP_T1,
    HDL_T0, HDL_T1,
    LDL_T0, LDL_T1,
    nonHDL_T0, nonHDL_T1
  ) %>%
  tidyr::pivot_longer(
    cols = -c(study_id, age, sex, BMI_baseline, fasting_duration_measured, CRP_T0_z),
    names_to = c("outcome", "time"),
    names_pattern = "(.+)_T(0|1)",
    values_to = "value"
  ) %>%
  dplyr::mutate(time = paste0("T", time)) %>%
  tidyr::pivot_wider(names_from = time, values_from = value) %>%
  dplyr::filter(outcome %in% all_outcomes) %>%
  dplyr::mutate(
    T0_model = dplyr::case_when(
      outcome %in% c("Insulin", "HOMA") ~ safe_log(T0),
      TRUE ~ T0
    ),
    T1_model = dplyr::case_when(
      outcome %in% c("Insulin", "HOMA") ~ safe_log(T1),
      TRUE ~ T1
    )
  )

# -------------------------
# 4) Fit one ANCOVA per outcome
# -------------------------
fit_crp_ancova_one <- function(data, outcome_name) {
  df <- data %>%
    dplyr::filter(outcome == outcome_name) %>%
    dplyr::filter(
      !is.na(T0_model), !is.na(T1_model),
      !is.na(CRP_T0_z),
      !is.na(fasting_duration_measured),
      !is.na(BMI_baseline),
      !is.na(age),
      !is.na(sex)
    ) %>%
    dplyr::mutate(
      T0_c      = center_only(T0_model),
      fasting_c = center_only(fasting_duration_measured),
      BMI_c     = center_only(BMI_baseline),
      age_c     = center_only(age),
      sex       = factor(sex, levels = c("F", "M"))
    )

  n_subj <- dplyr::n_distinct(df$study_id)

  if (n_subj < 30) {
    return(tibble(
      outcome = outcome_name,
      estimate = NA_real_,
      std.error = NA_real_,
      conf.low = NA_real_,
      conf.high = NA_real_,
      p.value = NA_real_,
      n = n_subj
    ))
  }

  fit <- lm(
    T1_model ~ T0_c + fasting_c + BMI_c + age_c + sex + CRP_T0_z,
    data = df
  )

  broom::tidy(fit, conf.int = TRUE) %>%
    dplyr::filter(term == "CRP_T0_z") %>%
    dplyr::transmute(
      outcome = outcome_name,
      estimate,
      std.error,
      conf.low,
      conf.high,
      p.value,
      n = n_subj
    )
}

# -------------------------
# 5) Run models for all outcomes
# -------------------------
main_effect_results_crp <- purrr::map_dfr(
  all_outcomes,
  ~ fit_crp_ancova_one(ancova_long_crp, .x)
) %>%
  dplyr::mutate(
    p_fdr = p.adjust(p.value, method = "BH")
  ) %>%
  dplyr::arrange(p_fdr)

print(main_effect_results_crp)

# Quick sanity checks
print(unique(main_effect_results_crp$outcome))
print(dplyr::count(ancova_long_crp, outcome))

# -------------------------
# 6) Pretty summary table
# -------------------------
crp_table_pretty <- main_effect_results_crp %>%
  dplyr::mutate(
    Outcome = dplyr::recode(
      outcome,
      Glucose = "Glucose",
      Insulin = "Insulin",
      HOMA = "HOMA-IR",
      TyG = "TyG index",
      AIP = "AIP",
      HDL = "HDL-C",
      LDL = "LDL-C",
      nonHDL = "Non-HDL-C"
    ),
    Estimate = round(estimate, 3),
    SE = round(std.error, 3),
    `95% CI` = ifelse(
      is.na(conf.low) | is.na(conf.high),
      NA_character_,
      paste0("[", round(conf.low, 3), ", ", round(conf.high, 3), "]")
    ),
    `P (raw)` = fmt_p(p.value),
    `P (FDR)` = fmt_p(p_fdr),
    `Signif.` = p_stars(p_fdr)
  ) %>%
  dplyr::arrange(p_fdr, p.value) %>%
  dplyr::select(
    Outcome, Estimate, SE, `95% CI`, `P (raw)`, `P (FDR)`, `Signif.`, n
  )

print(crp_table_pretty)

crp_table_pretty %>%
  kableExtra::kbl(
    caption = "ANCOVA results for baseline CRP predicting fasting-endpoint metabolic outcomes (T1), adjusted for baseline outcome level, fasting duration, BMI, age, and sex.",
    align = c("l", "r", "r", "l", "r", "r", "c", "r"),
    booktabs = TRUE
  ) %>%
  kableExtra::kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width = FALSE,
    font_size = 12
  ) %>%
  kableExtra::add_header_above(c(" " = 1, "CRP effect" = 6, " " = 1)) %>%
  kableExtra::footnote(
    general = "Estimate refers to the adjusted regression coefficient for baseline CRP (z-scored log-CRP).",
    general_title = "Note: "
  )

# -------------------------
# 7) Save CSV
# -------------------------
dir.create("ModelTables", showWarnings = FALSE)

readr::write_csv(
  crp_table_pretty,
  "ModelTables/ANCOVA_baseline_CRP_predicting_T1_metabolic_outcomes.csv"
)
```

## 15) CRP (T0->T1)

```{r CRp check, message=FALSE, warning=FALSE}
## ============================================================
## Baseline CRP stratification table
## Mirrors Mesnage et al. 2025 stratification approach
## Groups: Low (<1 mg/L), Moderate (1-10 mg/L), Elevated (>10 mg/L)
## ============================================================

suppressPackageStartupMessages({
  library(dplyr)
  library(tidyr)
  library(knitr)
  library(kableExtra)
})

## -------------------------
## 1) Build CRP wide (T0, T1)
## -------------------------
crp_wide <- FF_laboratory_master %>%
  filter(
    variable == "CRP (mg/l)",
    timepoint %in% c("T0", "T1")
  ) %>%
  group_by(study_id, timepoint) %>%
  slice_min(datetime, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  select(study_id, timepoint, value, BMI_baseline) %>%
  pivot_wider(names_from = timepoint, values_from = value) %>%
  filter(!is.na(T0), !is.na(T1), T0 > 0, T1 > 0) %>%
  mutate(
    delta_CRP     = T1 - T0,
    pct_change    = ((T1 - T0) / T0) * 100,
    CRP_baseline_group = case_when(
      T0 < 1  ~ "Low (<1 mg/L)",
      T0 <= 10 ~ "Moderate (1–10 mg/L)",
      T0 > 10  ~ "Elevated (>10 mg/L)"
    ),
    CRP_baseline_group = factor(
      CRP_baseline_group,
      levels = c("Low (<1 mg/L)", "Moderate (1–10 mg/L)", "Elevated (>10 mg/L)")
    ),
    BMI_cat = factor(
      case_when(
        is.na(BMI_baseline) ~ NA_character_,
        BMI_baseline < 25   ~ "Normal",
        BMI_baseline < 30   ~ "Overweight",
        TRUE                ~ "Obese"
      ),
      levels = c("Normal", "Overweight", "Obese")
    )
  )

## -------------------------
## 2) Summary table by CRP group
## -------------------------
fmt_mean_se <- function(x) {
  x <- x[!is.na(x)]
  if (length(x) == 0) return(NA_character_)
  paste0(round(mean(x), 2), " ± ", round(sd(x) / sqrt(length(x)), 2))
}

crp_strat_table <- crp_wide %>%
  group_by(CRP_baseline_group) %>%
  summarise(
    n                    = n(),
    `CRP at T0 (mean ± SE)`  = fmt_mean_se(T0),
    `CRP at T1 (mean ± SE)`  = fmt_mean_se(T1),
    `Absolute change (mean ± SE)` = fmt_mean_se(delta_CRP),
    `% change (mean ± SE)`   = fmt_mean_se(pct_change),
    .groups = "drop"
  )

## -------------------------
## 3) Add a total row
## -------------------------
total_row <- crp_wide %>%
  summarise(
    CRP_baseline_group   = factor("Total", levels = levels(crp_wide$CRP_baseline_group)),
    n                    = n(),
    `CRP at T0 (mean ± SE)`  = fmt_mean_se(T0),
    `CRP at T1 (mean ± SE)`  = fmt_mean_se(T1),
    `Absolute change (mean ± SE)` = fmt_mean_se(delta_CRP),
    `% change (mean ± SE)`   = fmt_mean_se(pct_change)
  )

crp_strat_final <- bind_rows(crp_strat_table, total_row) %>%
  rename(`Baseline CRP group` = CRP_baseline_group)

## -------------------------
## 4) Print table
## -------------------------
crp_strat_final %>%
  kbl(
    caption = "Table X. Fasting-associated changes in CRP stratified by baseline CRP level. 
               Values are mean ± SE. Low: <1 mg/L; Moderate: 1–10 mg/L; Elevated: >10 mg/L.",
    align = c("l", "c", "c", "c", "c", "c")
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width = FALSE
  ) %>%
  row_spec(nrow(crp_strat_final), bold = TRUE) %>%
  footnote(
    general = "CRP measured in mg/L. T0 = pre-fasting baseline; T1 = end of fasting endpoint.",
    general_title = "Note: "
  )

## -------------------------
## 5) Optional: also cross-tabulate by BMI category
## -------------------------
crp_bmi_table <- crp_wide %>%
  filter(!is.na(BMI_cat)) %>%
  group_by(CRP_baseline_group, BMI_cat) %>%
  summarise(
    n = n(),
    `Mean T0 CRP` = round(mean(T0), 2),
    `Mean T1 CRP` = round(mean(T1), 2),
    `Mean Δ CRP`  = round(mean(delta_CRP), 2),
    `Mean % change` = round(mean(pct_change), 1),
    .groups = "drop"
  ) %>%
  rename(
    `Baseline CRP group` = CRP_baseline_group,
    `BMI category` = BMI_cat
  )

crp_bmi_table %>%
  kbl(
    caption = "Table X (supplementary). Fasting-associated CRP changes by baseline CRP group and BMI category.",
    align = "l"
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width = FALSE
  ) %>%
  collapse_rows(columns = 1, valign = "top")

## -------------------------
## 6) Save both
## -------------------------
dir.create("ModelTables", showWarnings = FALSE)
write.csv(crp_strat_final,  "ModelTables/CRP_stratified_by_baseline_group.csv",     row.names = FALSE)
write.csv(crp_bmi_table,    "ModelTables/CRP_stratified_by_baseline_and_BMI.csv",   row.names = FALSE)
```

## 16) TG/HDL (T0->T1)

```{r TG/HDL check, message=FALSE, warning=FALSE}
## ============================================================
## TG/HDL ratio by baseline CRP stratum and BMI category
## At T0, T1, and change during fasting
## ============================================================

suppressPackageStartupMessages({
  library(dplyr)
  library(tidyr)
  library(knitr)
  library(kableExtra)
})

## -------------------------
## 1) Build TG/HDL wide (T0, T1)
## -------------------------
tg_hdl_wide <- FF_laboratory_master %>%
  filter(
    variable %in% c("Triglycerides (mg/dl)", "HDL-C (mg/dl)"),
    timepoint %in% c("T0", "T1")
  ) %>%
  group_by(study_id, variable, timepoint) %>%
  slice_min(datetime, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  select(study_id, variable, timepoint, value, BMI_baseline) %>%
  pivot_wider(names_from = c(variable, timepoint), values_from = value) %>%
  rename(
    TG_T0  = `Triglycerides (mg/dl)_T0`,
    TG_T1  = `Triglycerides (mg/dl)_T1`,
    HDL_T0 = `HDL-C (mg/dl)_T0`,
    HDL_T1 = `HDL-C (mg/dl)_T1`
  ) %>%
  filter(
    !is.na(TG_T0), !is.na(HDL_T0), HDL_T0 > 0,
    !is.na(TG_T1), !is.na(HDL_T1), HDL_T1 > 0
  ) %>%
  mutate(
    TG_HDL_T0 = TG_T0 / HDL_T0,
    TG_HDL_T1 = TG_T1 / HDL_T1,
    delta_TG_HDL = TG_HDL_T1 - TG_HDL_T0,
    pct_change   = ((TG_HDL_T1 - TG_HDL_T0) / TG_HDL_T0) * 100,
    BMI_cat = factor(
      case_when(
        is.na(BMI_baseline) ~ NA_character_,
        BMI_baseline < 25   ~ "Normal",
        BMI_baseline < 30   ~ "Overweight",
        TRUE                ~ "Obese"
      ),
      levels = c("Normal", "Overweight", "Obese")
    )
  )

## -------------------------
## 2) Join baseline CRP groups
## -------------------------
crp_groups <- FF_laboratory_master %>%
  filter(variable == "CRP (mg/l)", timepoint == "T0") %>%
  group_by(study_id) %>%
  slice_min(datetime, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  transmute(
    study_id,
    CRP_T0 = value,
    CRP_baseline_group = factor(
      case_when(
        value < 1   ~ "Low (<1 mg/L)",
        value <= 10 ~ "Moderate (1–10 mg/L)",
        value > 10  ~ "Elevated (>10 mg/L)"
      ),
      levels = c("Low (<1 mg/L)", "Moderate (1–10 mg/L)", "Elevated (>10 mg/L)")
    )
  )

tg_hdl_crp <- tg_hdl_wide %>%
  left_join(crp_groups, by = "study_id") %>%
  filter(!is.na(CRP_baseline_group))

## -------------------------
## 3) Summary by CRP stratum only
## -------------------------
fmt_mean_sd <- function(x) {
  x <- x[!is.na(x)]
  paste0(round(mean(x), 2), " ± ", round(sd(x), 2))
}

tg_hdl_by_crp <- tg_hdl_crp %>%
  group_by(CRP_baseline_group) %>%
  summarise(
    n                        = n(),
    `TG/HDL at T0 (mean ± SD)` = fmt_mean_sd(TG_HDL_T0),
    `TG/HDL at T1 (mean ± SD)` = fmt_mean_sd(TG_HDL_T1),
    `Δ TG/HDL (mean ± SD)`     = fmt_mean_sd(delta_TG_HDL),
    `% change (mean ± SD)`     = fmt_mean_sd(pct_change),
    .groups = "drop"
  )

## -------------------------
## 4) Summary by CRP stratum AND BMI category
## -------------------------
tg_hdl_by_crp_bmi <- tg_hdl_crp %>%
  filter(!is.na(BMI_cat)) %>%
  group_by(CRP_baseline_group, BMI_cat) %>%
  summarise(
    n                  = n(),
    `Mean T0 TG/HDL`   = round(mean(TG_HDL_T0, na.rm = TRUE), 2),
    `Mean T1 TG/HDL`   = round(mean(TG_HDL_T1, na.rm = TRUE), 2),
    `Mean Δ TG/HDL`    = round(mean(delta_TG_HDL, na.rm = TRUE), 2),
    `Mean % change`    = round(mean(pct_change, na.rm = TRUE), 1),
    .groups = "drop"
  ) %>%
  rename(
    `Baseline CRP group` = CRP_baseline_group,
    `BMI category`       = BMI_cat
  )

## -------------------------
## 5) Print tables
## -------------------------
tg_hdl_by_crp %>%
  rename(`Baseline CRP group` = CRP_baseline_group) %>%
  kbl(
    caption = "TG/HDL ratio by baseline CRP stratum (T0 and T1).",
    align = c("l", "c", "c", "c", "c", "c")
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width = FALSE
  )

tg_hdl_by_crp_bmi %>%
  kbl(
    caption = "TG/HDL ratio by baseline CRP stratum and BMI category.",
    align = "l"
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width = FALSE
  ) %>%
  collapse_rows(columns = 1, valign = "top")

## -------------------------
## 6) Save
## -------------------------
dir.create("ModelTables", showWarnings = FALSE)
write.csv(tg_hdl_by_crp,
          "ModelTables/TGHDL_by_CRP_stratum.csv",
          row.names = FALSE)
write.csv(tg_hdl_by_crp_bmi,
          "ModelTables/TGHDL_by_CRP_stratum_and_BMI.csv",
          row.names = FALSE)
```

## 17) Residuals and QQ plotsfor main ANCOVA models

```{r ANCOVA-main-diagnostics, message=FALSE, warning=FALSE, fig.width=16, fig.height=10}
## ============================================================
## ANCOVA Residual Diagnostics
## Fasting:   T1 ~ T0_c + fast_c + BMI_c + age_c + sex
## Refeeding: T2 ~ T1_c + diet_keto + FR_c + age_c + sex + BMI_c
##
## Markers: CRP, GPT/ALT, Insulin, Glucose, Neutrophils,
##          Monocytes, WBC, Lymphocytes, LDL-C, Total Cholesterol
## Output:  one panel (resid + QQ) per phase
## ============================================================

diag_markers_select <- c(
  "CRP (mg/l)",
  "GPT/ALT (U/l)",
  "Insulin (mIU/l)",
  "Serum Glucose (mg/dl)",
  "Neutrophils abs. (Tsd./µl)",
  "Monocytes abs. (Tsd./µl)",
  "White Blood Cells (Tsd./µl)",
  "Lymphocytes abs. (Tsd./µl)",
  "LDL-C (mg/dl)",
  "Total Cholesterol (mg/dl)"
)

## ---- helper: fit one marker and return augmented residuals ----

get_resid_df <- function(wide_df, outcome_col, predictor_cols,
                         marker_name, log_markers = markers_log) {

  df <- wide_df %>%
    filter(variable == marker_name) %>%
    drop_na(all_of(c(outcome_col, predictor_cols)))

  if (n_distinct(df$study_id) < 20) return(NULL)

  fit <- lm(
    as.formula(paste(outcome_col, "~", paste(predictor_cols, collapse = " + "))),
    data = df
  )

  is_logged <- marker_name %in% log_markers   # explicit, not relying on global scope

  broom::augment(fit) %>%
    mutate(
      marker       = marker_name,
      display_name = ifelse(is_logged, paste0(marker_name, " (log)"), marker_name),
      label = display_name %>%
        str_replace(" abs\\. \\(Tsd\\..*?\\)", "") %>%
        str_replace(" \\(mg/dl\\)", "") %>%
        str_replace(" \\(U/l\\)", "") %>%
        str_replace(" \\(mIU/l\\)", "") %>%
        str_replace("Serum ", "") %>%
        str_replace("White Blood Cells", "WBC") %>%
        str_replace(" \\(log\\)", "\n(log)")
    )
}

## Then call with explicit log_markers argument
fast_resid <- purrr::map_dfr(
  diag_markers_select,
  ~ get_resid_df(
      wide_df        = fasting_wide,
      outcome_col    = "T1",
      predictor_cols = c("T0_c", "fast_c", "BMI_c", "age_c", "sex"),
      marker_name    = .x,
      log_markers    = markers_log   # explicit
    )
)

refeed_resid <- purrr::map_dfr(
  diag_markers_select,
  ~ get_resid_df(
      wide_df        = refeed_wide,
      outcome_col    = "T2",
      predictor_cols = c("T1_c", "diet_keto", "FR_c", "age_c", "sex", "BMI_c"),
      marker_name    = .x,
      log_markers    = markers_log   # explicit
    )
)

## ---- fix factor order for panels ----
marker_label_order <- fast_resid %>%
  distinct(marker, label) %>%
  slice(match(diag_markers_select, marker)) %>%
  pull(label)

fast_resid  <- fast_resid  %>% mutate(label = factor(label, levels = marker_label_order))
refeed_resid <- refeed_resid %>% mutate(label = factor(label, levels = marker_label_order))

## ============================================================
## PLOT FUNCTION: residuals vs fitted + QQ side by side per marker
## ============================================================
make_diag_panel <- function(resid_df, phase_label) {

  # --- Residuals vs Fitted ---
  p_resid <- ggplot(resid_df, aes(x = .fitted, y = .resid)) +
    geom_hline(yintercept = 0, linetype = "dashed", colour = "grey50", linewidth = 0.6) +
    geom_point(alpha = 0.45, size = 0.9, colour = "grey30") +
    geom_smooth(
      method  = "loess", se = FALSE,
      colour  = "steelblue", linewidth = 0.85
    ) +
    facet_wrap(~ label, scales = "free", ncol = 5) +
    labs(
      title    = paste0(phase_label, " ANCOVA — Residuals vs Fitted"),
      subtitle = "Loess smoother should be flat; curvature = potential misspecification",
      x        = "Fitted values",
      y        = "Residuals"
    ) +
    theme_bw(base_size = 11) +
    theme(
      strip.background = element_rect(fill = "grey92"),
      strip.text       = element_text(face = "bold", size = 9),
      panel.grid.minor = element_blank(),
      plot.title       = element_text(face = "bold")
    )

  # --- QQ plot ---
  p_qq <- ggplot(resid_df, aes(sample = .std.resid)) +
    stat_qq(alpha = 0.45, size = 0.9, colour = "grey30") +
    stat_qq_line(colour = "steelblue", linewidth = 0.85) +
    facet_wrap(~ label, scales = "free", ncol = 5) +
    labs(
      title    = paste0(phase_label, " ANCOVA — Normal QQ Plot"),
      subtitle = "Points should follow the reference line",
      x        = "Theoretical quantiles",
      y        = "Standardised residuals"
    ) +
    theme_bw(base_size = 11) +
    theme(
      strip.background = element_rect(fill = "grey92"),
      strip.text       = element_text(face = "bold", size = 9),
      panel.grid.minor = element_blank(),
      plot.title       = element_text(face = "bold")
    )

  list(resid = p_resid, qq = p_qq)
}

## ---- generate panels ----
fast_plots   <- make_diag_panel(fast_resid,   "Fasting")
refeed_plots <- make_diag_panel(refeed_resid, "Refeeding")

## ---- print in document ----
print(fast_plots$resid)
print(fast_plots$qq)
print(refeed_plots$resid)
print(refeed_plots$qq)

## ---- save ----
dir.create("ModelDiagnostics", showWarnings = FALSE)

ggsave("ModelDiagnostics/Fasting_ANCOVA_resid_vs_fitted.png",
       fast_plots$resid,   width = 16, height = 7, dpi = 300)
ggsave("ModelDiagnostics/Fasting_ANCOVA_QQ.png",
       fast_plots$qq,      width = 16, height = 7, dpi = 300)
ggsave("ModelDiagnostics/Refeeding_ANCOVA_resid_vs_fitted.png",
       refeed_plots$resid, width = 16, height = 7, dpi = 300)
ggsave("ModelDiagnostics/Refeeding_ANCOVA_QQ.png",
       refeed_plots$qq,    width = 16, height = 7, dpi = 300)

## ---- companion summary table ----
diag_summary <- bind_rows(
  fast_resid   %>% mutate(phase = "Fasting"),
  refeed_resid %>% mutate(phase = "Refeeding")
) %>%
  group_by(phase, marker) %>%
  summarise(
    n             = n(),
    max_cook      = round(max(cooks.distance(
                      lm(.resid ~ .fitted)   # placeholder — Cook's already in augment
                    ), na.rm = TRUE), 4),
    loess_max_dev = {
      lo <- loess(.resid ~ .fitted, data = cur_data())
      round(max(abs(predict(lo)), na.rm = TRUE), 3)
    },
    .groups = "drop"
  ) %>%
  mutate(
    log_transformed = marker %in% markers_log,
    flag = ifelse(loess_max_dev > 0.3, "⚠ Review", "OK")
  ) %>%
  arrange(phase, desc(loess_max_dev))

diag_summary %>%
  kbl(
    caption = paste0(
      "Table S1. Residual diagnostic summary. loess_max_dev = maximum absolute deviation of Loess ",
      "smoother from zero (proxy for non-linearity/heteroscedasticity). ",
      "† markers were log-transformed prior to modelling."
    ),
    align = "l"
  ) %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)

write_csv(diag_summary, "ModelTables/ANCOVA_residual_diagnostic_summary.csv")
```

## 18) Cook's distance and refitting 

```{r Cooksdistanceandrefitting, message=FALSE, warning=FALSE, fig.width=16, fig.height=10}
# due to outliers in some of the markers, especially even after log-transformation, the observation with the most influence was removed and the model was refitted again.
# model was defined as robust if SE change < 1
# since it takes SE change = 2 to go from significant to ns or vice versa 

run_cooks_sensitivity <- function(wide_df, outcome, predictors, 
                                   key_predictors, phase_label) {
  wide_df %>%
    group_by(variable) %>%
    group_modify(~{
      df <- .x %>% drop_na(all_of(c(outcome, predictors)))
      
      n_subj <- n_distinct(df$study_id)
      if (n_subj < 20) return(tibble(
        n = n_subj, max_cook = NA_real_,
        worst_subject = NA_character_,
        max_change_SE = NA_real_,
        conclusion = "Skipped"
      ))
      
      fit_full <- lm(
        as.formula(paste(outcome, "~", paste(predictors, collapse = "+"))),
        data = df
      )
      
      cooks     <- cooks.distance(fit_full)
      worst_idx <- which.max(cooks)
      
      df_reduced  <- df[-worst_idx, ]
      fit_reduced <- lm(
        as.formula(paste(outcome, "~", paste(predictors, collapse = "+"))),
        data = df_reduced
      )
      
      coef_full    <- coef(fit_full)[key_predictors]
      coef_reduced <- coef(fit_reduced)[key_predictors]
      se_full      <- summary(fit_full)$coefficients[key_predictors, "Std. Error"]
      
      max_change_in_se <- max(
        abs(coef_reduced - coef_full) / se_full,
        na.rm = TRUE
      )
      
      robust <- max_change_in_se < 1
      
      tibble(
        n             = n_subj,
        max_cook      = round(max(cooks), 4),
        worst_subject = as.character(df$study_id[worst_idx]),
        max_change_SE = round(max_change_in_se, 2),
        conclusion    = ifelse(
          robust,
          paste0("Robust (max ", round(max_change_in_se, 2), " SE change)"),
          paste0("⚠ Review (max ", round(max_change_in_se, 2), " SE change)")
        )
      )
    }) %>%
    ungroup() %>%
    mutate(phase = phase_label)
}

## ---- fasting: key predictors are fast_c and BMI_c ----
cooks_fast <- run_cooks_sensitivity(
  fasting_wide,
  outcome         = "T1",
  predictors      = c("T0_c", "fast_c", "BMI_c", "age_c", "sex"),
  key_predictors  = c("fast_c", "BMI_c"),
  phase_label     = "Fasting"
)

## ---- refeeding: key predictors are diet_keto and FR_c ----
cooks_refeed <- run_cooks_sensitivity(
  refeed_wide,
  outcome         = "T2",
  predictors      = c("T1_c", "diet_keto", "FR_c", "age_c", "sex", "BMI_c"),
  key_predictors  = c("diet_ketoKeto", "FR_c"),
  phase_label     = "Refeeding"
)

cooks_all <- bind_rows(cooks_fast, cooks_refeed) %>%
  arrange(phase, desc(max_cook))

cat(sprintf("Non-robust after removal: %d / %d\n",
  sum(grepl("⚠", cooks_all$conclusion), na.rm = TRUE),
  nrow(cooks_all)))

cooks_all %>%
  kbl(
    caption = paste0(
      "Table S2. Cook's distance sensitivity analysis. The single most influential observation ",
      "per marker was removed and key predictors refit. Robustness assessed as ",
      "change < 1 SE in key predictor coefficients after removal."
    ),
    align = "l"
  ) %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE) %>%
  scroll_box(width = "100%", height = "500px")

write_csv(cooks_all, "ModelTables/Cooks_sensitivity_main_ANCOVA.csv")
```

## 19) VIF ANCOVA tests

```{r vif-all-ancova, message=FALSE, warning=FALSE}
## ============================================================
## S1.3  VIF — ALL fasting + refeeding ANCOVA models
## ============================================================

suppressPackageStartupMessages(library(car))

# Helper: fit one marker and return VIF
get_vif <- function(wide_df, outcome, predictors, phase_label) {
  wide_df %>%
    group_by(variable) %>%
    group_modify(~{
      df <- .x %>% drop_na(all_of(c(outcome, predictors)))
      
      if (n_distinct(df$study_id) < 20) return(tibble(predictor = predictors, VIF = NA_real_))
      
      fit      <- lm(as.formula(paste(outcome, "~", paste(predictors, collapse = "+"))), data = df)
      vif_vals <- tryCatch(car::vif(fit), error = function(e) rep(NA_real_, length(predictors)))
      
      tibble(predictor = names(vif_vals), VIF = round(as.numeric(vif_vals), 2))
    }) %>%
    ungroup() %>%
    mutate(phase = phase_label)
}

# Fasting: T1 ~ T0_c + fast_c + BMI_c + age_c + sex
vif_fast <- get_vif(
  fasting_wide,
  outcome    = "T1",
  predictors = c("T0_c", "fast_c", "BMI_c", "age_c", "sex"),
  phase_label = "Fasting"
)

# Refeeding: T2 ~ T1_c + diet_keto + FR_c + age_c + sex + BMI_c
vif_refeed <- get_vif(
  refeed_wide,
  outcome    = "T2",
  predictors = c("T1_c", "diet_keto", "FR_c", "age_c", "sex", "BMI_c"),
  phase_label = "Refeeding"
)

vif_all <- bind_rows(vif_fast, vif_refeed) %>%
  mutate(flag = case_when(
    is.na(VIF)  ~ "–",
    VIF >= 10   ~ "⚠ Severe (≥10)",
    VIF >=  5   ~ "⚠ Moderate (≥5)",
    TRUE        ~ "OK"
  ))

# Max VIF per marker — clean summary
vif_summary <- vif_all %>%
  filter(!is.na(VIF)) %>%
  group_by(phase, variable) %>%
  summarise(
    max_VIF    = max(VIF),
    worst_pred = predictor[which.max(VIF)],
    .groups = "drop"
  ) %>%
  mutate(flag = case_when(
    max_VIF >= 10 ~ "⚠ Severe",
    max_VIF >=  5 ~ "⚠ Moderate",
    TRUE          ~ "OK"
  )) %>%
  arrange(phase, desc(max_VIF))

cat(sprintf("Max VIF across all models: %.2f\n", max(vif_summary$max_VIF, na.rm = TRUE)))
cat(sprintf("Models with VIF ≥ 5: %d\n", sum(vif_summary$max_VIF >= 5, na.rm = TRUE)))

vif_summary %>%
  kbl(
    caption = paste0(
      "Table S3. Maximum VIF per marker for fasting and refeeding ANCOVA models. ",
      "All predictors were mean-centred prior to modelling. ",
      "VIF < 5 indicates no meaningful collinearity."
    ),
    align = "l"
  ) %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE) %>%
  scroll_box(width = "100%", height = "500px")

write_csv(vif_summary, "ModelTables/S1_VIF_all_ANCOVA.csv")
```


## 20) Coefficient of variation

```{r CV-all, message=FALSE, warning=FALSE}
# CV is dimensionless so dont worry about units
# preferably use raw and untransformed values
# mean == phase specific mean change
# abs() afterwards since direction doesnt matter for change

delta_df <- FF_laboratory_master %>%
  filter(variable %in% markers_lm, timepoint %in% c("T0", "T1", "T2")) %>%
  group_by(study_id, variable, timepoint) %>%
  slice_min(datetime, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  pivot_wider(
    id_cols    = c(study_id, sex, variable, fasting_duration_measured,
                   FR_duration, age, BMI_baseline, diet_keto),
    names_from  = timepoint,
    values_from = value
  ) %>%
  mutate(
    delta_T1_T0 = if_else(!is.na(T1) & !is.na(T0), T1 - T0, NA_real_),
    delta_T2_T1 = if_else(!is.na(T2) & !is.na(T1), T2 - T1, NA_real_)
  )

cv_fasting <- delta_df %>%
  mutate(domain = assign_domain_from_variable(variable)) %>%
  filter(domain %in% c("Circulating immune cells", "Lipid", "Metabolic / Organ")) %>%
  group_by(domain, variable) %>%
  summarise(
    mean = mean(delta_T1_T0, na.rm = TRUE),
    sd   = sd(delta_T1_T0,   na.rm = TRUE),
    CV   = abs(sd / mean) * 100,
    .groups = "drop"
  ) %>%
  group_by(domain) %>%
  summarise(median_CV_fasting = median(CV, na.rm = TRUE), .groups = "drop")

cv_refeeding <- delta_df %>%
  mutate(domain = assign_domain_from_variable(variable)) %>%
  filter(domain %in% c("Circulating immune cells", "Lipid", "Metabolic / Organ")) %>%
  group_by(domain, variable) %>%
  summarise(
    mean = mean(delta_T2_T1, na.rm = TRUE),
    sd   = sd(delta_T2_T1,   na.rm = TRUE),
    CV   = abs(sd / mean) * 100,
    .groups = "drop"
  ) %>%
  group_by(domain) %>%
  summarise(median_CV_refeeding = median(CV, na.rm = TRUE), .groups = "drop")

left_join(cv_fasting, cv_refeeding, by = "domain") %>%
  mutate(
    median_CV_fasting   = round(median_CV_fasting,   1),
    median_CV_refeeding = round(median_CV_refeeding, 1)
  ) %>%
  arrange(desc(median_CV_fasting)) %>%
  rename(
    "Domain"             = domain,
    "Fasting (T0→T1)"   = median_CV_fasting,
    "Refeeding (T1→T2)" = median_CV_refeeding
  ) %>%
  kbl(
    caption = "Table S. Median coefficient of variation (%) of within-person changes by domain and phase.",
    align   = c("l", "c", "c")
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width = FALSE
  ) %>%
  add_header_above(c(" " = 1, "Median CV (%)" = 2)) %>%
  footnote(
    general       = "CV = |SD / mean| × 100, computed on raw untransformed within-person changes. Higher values indicate greater inter-individual variability relative to the average change.",
    general_title = "Note: "
  )
```